### 2024-07-24 tex/cover_latter.tex
after resubmission

\documentclass{letter}
\usepackage{hyperref}
\usepackage{fullpage}

\signature{Dr Geraint Palmer}
\begin{document}

\begin{letter}{}
\opening{Dear editors,}

My name is Dr Geraint Palmer, lecturer at the School of Mathematics at Cardiff University. I am writing on behalf on of a team of authors as we wish to submit a research article entitled \textit{``Optimising Heterogeneous Ambulance Fleet Allocations in Jakarta''} for consideration in the European Journal of Operational Research.

In this paper, we propose a modelling approach to consider where to optimally locate different types of emergency response vehicles in order to maximise patient outcomes within a heterogeneous population. We use this modelling approach within a heuristic algorithm to optimally allocate the vehicles, and use a novel discrete event simulation to evaluate the found allocations.
This paper is a continuation and extension of previously published work, namely \href{https://www.sciencedirect.com/science/article/pii/S0305048312000436?casa_token=1SiRonMwOYIAAAAA:AnZOLnmzNdYATlXDMYjqnQHEZX701o1eaBypnyMpgQbqoR62Z57aX1SXAadw4GJAku8gNvc}{\textit{``Ambulance allocation for maximal survival with heterogeneous outcome measures''}} by Knight, Harper and Smith.
This paper extends that work of that paper in four ways:

\begin{itemize}
  \item we extend the model to include heterogeneous fleets, that is both primary transport vehicles and secondary response vehicles;
  \item we overcome one problem identified in the first paper, that is the circular nature of the vehicle utilisations, by numerically solving implicit equations relating allocation and utilisation;
  \item we develop a sequential discrete event simulation of the calls and emergency response vehicles' logic to evaluate the found allocations on a range of KPIs;
  \item we apply the work to the real life situation in the city of Jakarta, in a developing country.
\end{itemize}

Thank you for your consideration of this article.

\closing{Yours Faithfully,}

\end{letter}
\end{document}

### 2024-07-24 tex/ima-authoring-template.tex
after resubmission

%%
%% Copyright 2021 OXFORD UNIVERSITY PRESS
%%
%% This file is part of the 'ima-authoring-template Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'ima-authoring-template Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for OXFORD UNIVERSITY PRESS's document class `ima-authoring-template'
%% with bibliographic references
%%

\documentclass[numbers,webpdf,imaiai]{ima-authoring-template}%
%\documentclass[namedate,webpdf,imaman]{ima-authoring-template}

%\usepackage{showframe}

\graphicspath{{Fig/}}

% line numbers
%\usepackage[mathlines, switch]{lineno}
%\usepackage[right]{lineno}

\theoremstyle{thmstyletwo}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}%
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.

\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%

\newtheorem{definition}{Definition}

\numberwithin{equation}{section}

\begin{document}

\DOI{DOI HERE}
\copyrightyear{2021}
\vol{00}
\pubyear{2021}
\access{Advance Access Publication Date: Day Month Year}
\appnotes{Paper}
\copyrightstatement{Published by Oxford University Press on behalf of the Institute of Mathematics and its Applications. All rights reserved.}
\firstpage{1}

%\subtitle{Subject Section}

\title[Short Article Title]{Article Title}

\author{First Author*
\address{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}}
\author{Second Author and Third Author\ORCID{0000-0000-0000-0000}
\address{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}}
\author{Fourth Author
\address{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}}
\author{Fifth Author
\address{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}}

\authormark{Author Name et al.}

\corresp[*]{Corresponding author: \href{email:email-id.com}{email-id.com}}

\received{Date}{0}{Year}
\revised{Date}{0}{Year}
\accepted{Date}{0}{Year}

%\editor{Associate Editor: Name}

\abstract{Abstracts must be able to stand alone and so cannot contain citations to
the paper's references, equations, etc. An abstract must consist of a single
paragraph and be concise. Because of online formatting, abstracts must appear
as plain as possible.}
\keywords{keyword1; Keyword2; Keyword3; Keyword4.}

% \boxedtext{
% \begin{itemize}
% \item Key boxed text here.
% \item Key boxed text here.
% \item Key boxed text here.
% \end{itemize}}

\maketitle

\section{Introduction}
The introduction introduces the context and summarizes the manuscript. It is importantly to clearly state the contributions
of this piece of work.

Nam dui ligula, fringilla a, euismod sodales, sollicitudin\footnote{\url{https://data.gov.uk/}} vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque\footnote{\url{https://www.ukdataservice.ac.uk/}}
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat\footnote{That is not to say that there have been no breaches---researchers using data for a project they are not licensed for or passing data
to unlicensed colleagues---however, this, as far as we are aware of, has never caused harm to individuals or entities in the data.} ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

\section{This is an example for first level head - section head}\label{sec2}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris (refer Section~\ref{sec5}).

\subsection{This is an example for second level head - subsection head}\label{subsec1}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.
\subsubsection{This is an example for third level head - subsubsection head}\label{subsubsec1}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

\paragraph{This is an example for fourth level head - paragraph head}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante.

\section{This is an example for first level head}\label{sec3}

\subsection{This is an example for second level head - subsection head}\label{subsec2}

\subsubsection{This is an example for third level head - subsubsection head}\label{subsubsec2}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi.

\paragraph{This is an example for fourth level head - paragraph head}

 Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

\section{Equations}\label{sec4}

Equations in \LaTeX{} can either be inline or on-a-line by itself. For
inline equations use the \verb+$...$+ commands. Eg: The equation
$H\psi = E \psi$ is written via the command \verb+$H \psi = E \psi$+.

For on-a-line by itself equations (with auto generated equation numbers)
one can use the equation environment:
\begin{equation}
\|\tilde{X}(k)\|^2 \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q}.\label{eq1}
\end{equation}
where,
\begin{align}
D_\mu &=  \partial_\mu - ig \frac{\lambda^a}{2} A^a_\mu, \nonumber \\
F^a_{\mu\nu} &= \partial_\mu A^a_\nu - \partial_\nu A^a_\mu + g f^{abc} A^b_\mu A^a_\nu.\label{eq2}
\end{align}
Notice the use of \verb+\nonumber+ in the align environment at the end
of each line, except the last, so as not to produce equation numbers on
lines where no equation numbers are required. The \verb+\label{}+ command
should only be used at the last line of an align environment where
\verb+\nonumber+ is not used.
\begin{equation}
Y_\infty = \left( \frac{m}{\textrm{GeV}} \right)^{-3}
    \left[ 1 + \frac{3 \ln(m/\textrm{GeV})}{15}
    + \frac{\ln(c_2/5)}{15} \right]
\end{equation}
The class file also supports the use of \verb+\mathbb{}+, \verb+\mathscr{}+ and
\verb+\mathcal{}+ commands. As such \verb+\mathbb{R}+, \verb+\mathscr{R}+
and \verb+\mathcal{R}+ produces $\mathbb{R}$, $\mathscr{R}$ and $\mathcal{R}$
respectively (refer Subsubsection~\ref{subsubsec3}).

\section{Tables}\label{sec5}

Tables can be inserted via the normal table and tabular environment. To put
footnotes inside tables one has to 12pt use the additional ``tablenotes" environment
enclosing the tabular environment. The footnote appears just below the table
itself (refer Tables~\ref{tab1} and \ref{tab2}).

\begin{verbatim}
\begin{table}[t]
\caption{<table-caption>\label{<table-label>}}%
\begin{tabular}{@{}llll@{}}
\toprule
column 1 & column 2 & column 3 & column 4\\
\midrule
row 1 & data 1 & data 2          & data 3 \\
row 2 & data 4 & data 5$^{1}$ & data 6 \\
row 3 & data 7 & data 8      & data 9$^{2}$\\
\botrule
\end{tabular}
\begin{tablenotes}%
\item Source: Example for source.
\item[$^{1}$] Example for a 1st table footnote.
\item[$^{2}$] Example for a 2nd table footnote.
\end{tablenotes}
\end{table}
\end{verbatim}

Lengthy tables which doesn't fit in textwidth should be set as rotated table. For this, we need to use \verb+\begin{sidewaystable}...+ \verb+\end{sidewaystable}+ instead of\break \verb+\begin{table}...+ \verb+\end{table}+ environment.

\begin{table}[!t]
\caption{Caption text\label{tab1}}%
\begin{tabular*}{\columnwidth}{@{\extracolsep\fill}llll@{\extracolsep\fill}}
\toprule
column 1 & column 2  & column 3 & column 4\\
\midrule
row 1    & data 1   & data 2  & data 3  \\
row 2    & data 4   & data 5$^{1}$  & data 6  \\
row 3    & data 7   & data 8  & data 9$^{2}$  \\
\botrule
\end{tabular*}
\begin{tablenotes}%
\item Source: This is an example of table footnote this is an example of table footnote this is an example of table footnote this is an example of~table footnote this is an example of table footnote
\item[$^{1}$] Example for a first table footnote.
\item[$^{2}$] Example for a second table footnote.
\end{tablenotes}
\end{table}

\begin{table*}[h]
\caption{Example of a lengthy table which is set to full textwidth.\label{tab2}}
\tabcolsep=0pt%%
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lcccccc@{\extracolsep{\fill}}}
\toprule%
& \multicolumn{3}{@{}c@{}}{Element 1$^{1}$} & \multicolumn{3}{@{}c@{}}{Element 2$^{2}$} \\
\cline{2-4}\cline{5-7}%
Project & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ \\
\midrule
Element 3  & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$\\
Element 4  & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$\\
\botrule
\end{tabular*}
\begin{tablenotes}%
\item Note: This is an example of table footnote this is an example of table footnote this is an example of table footnote this is an example of~table footnote this is an example of table footnote
\item[$^{1}$] Example for a first table footnote.
\item[$^{2}$] Example for a second table footnote.\vspace*{-12pt}
\end{tablenotes}
\end{table*}

\begin{sidewaystable}%[!p]
\caption{Tables which are too long to fit, should be written using the ``sidewaystable" environment as shown here\label{tab3}}
\tabcolsep=0pt%
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lcccccc@{\extracolsep{\fill}}}
\toprule%
& \multicolumn{3}{@{}c@{}}{Element 1$^{1}$}& \multicolumn{3}{@{}c@{}}{Element$^{2}$} \\
\cline{2-4}\cline{5-7}%
Projectile & Energy     & $\sigma_{calc}$ & $\sigma_{expt}$ & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ \\
\midrule
Element 3 & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$ \\
Element 4 & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$ \\
Element 3 & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$ \\
Element 4 & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$ \\
Element 3 & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$ \\
Element 4 & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$ \\
Element 3 & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$ \\
Element 4 & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$ \\
Element 3 & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$ \\
Element 4 & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$ \\
Element 3 & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$ \\
Element 4 & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$ \\
\botrule
\end{tabular*}
\begin{tablenotes}%
\item Note: This is an example of table footnote this is an example of table footnote this is an example of table footnote this is an example of~table footnote this is an example of table footnote
\item[$^{1}$] This is an example of table footnote
\end{tablenotes}
\end{sidewaystable}

\section{Figures}\label{sec6}

As per the \LaTeX\ standards one has to use eps images for \verb+latex+ compilation and \verb+pdf/jpg/png+ images for
\verb+pdflatex+ compilation. This is one of the major difference between \verb+latex+
and \verb+pdflatex+. The images should be single page documents. The command for inserting images
for \verb+latex+ and \verb+pdflatex+ can be generalized. The package used to insert images in \verb+latex/pdflatex+ is the
graphicx package. Figures can be inserted via the normal figure environment as shown in the below example:

\begin{figure}[!t]%
\centering
\includegraphics{01}
\caption{This is a widefig. This is an example of long caption this is an example of long caption  this is an example of long caption this is an example of long caption}\label{fig1}
\end{figure}

\begin{figure*}[!t]%
\centering
\includegraphics{02}
\caption{This is a widefig. This is an example of long caption this is an example of long caption  this is an example of long caption this is an example of long caption}\label{fig2}
\end{figure*}

\begin{verbatim}
\begin{figure}[!t]
        \centering\includegraphics{<eps-file>}
        \caption{<figure-caption>}
        \label{<figure-label>}
\end{figure}
\end{verbatim}

For sample purpose, we have included the width of images in the optional argument of \verb+\includegraphics+ tag. Please ignore this. Lengthy figures which doesn't fit in textwidth should be set as rotated mode. For rotated figures, we need to use \verb+\begin{sidewaysfigure}+ \verb+...+ \verb+\end{sidewaysfigure}+ instead of \verb+\begin{figure}+ \verb+...+ \verb+\end{figure}+ environment.

\begin{sidewaysfigure}[!p]%
\centering
\includegraphics{01}
\caption{This is an example for sideways figure. This is an example of long caption this is an example of long caption  this is an example of long caption this is an example of long caption}\label{fig3}
\end{sidewaysfigure}

\section{Algorithms, Program codes and Listings}\label{sec7}

Packages \verb+algorithm+, \verb+algorithmicx+ and \verb+algpseudocode+ are used for setting algorithms in latex.
For this, one has to use the below format:

\begin{verbatim}
\begin{algorithm}
\caption{<alg-caption>}\label{<alg-label>}
\begin{algorithmic}[1]
. . .
\end{algorithmic}
\end{algorithm}
\end{verbatim}

We need to refer above listed package documentations for more details before setting \verb+algorithm+ environment.
To set program codes, one has to use \verb+program+, package. We need to use \verb+\begin{program}+ \verb+...+
\verb+\end{program}+ environment to set program codes. A fast exponentiation procedure:

\begin{algorithm}[!t]
\caption{Calculate $y = x^n$}\label{algo1}
\begin{algorithmic}[1]
\Require $n \geq 0 \vee x \neq 0$
\Ensure $y = x^n$
\State $y \Leftarrow 1$
\If{$n < 0$}
        \State $X \Leftarrow 1 / x$
        \State $N \Leftarrow -n$
\Else
        \State $X \Leftarrow x$
        \State $N \Leftarrow n$
\EndIf
\While{$N \neq 0$}
        \If{$N$ is even}
            \State $X \Leftarrow X \times X$
            \State $N \Leftarrow N / 2$
        \Else[$N$ is odd]
            \State $y \Leftarrow y \times X$
            \State $N \Leftarrow N - 1$
        \EndIf
\EndWhile
\end{algorithmic}
\end{algorithm}

Similarly, for \verb+listings+, one has to use \verb+listings+ package. \verb+\begin{lstlisting}+ \verb+...+ \verb+\end{lstlisting}+ environment is used to set environments similar to \verb+verbatim+ environment. Refer \verb+lstlisting+, package documentation for more details on this.

\begin{minipage}{\hsize}%
\lstset{language=Pascal}% Set your language (you can change the language for each code-block optionally)
\begin{lstlisting}[frame=single,framexleftmargin=-1pt,framexrightmargin=-17pt,framesep=12pt,linewidth=0.98\textwidth]
for i:=maxint to 0 do
begin
{ do nothing }
end;
Write('Case insensitive ');
Write('Pascal keywords.');
\end{lstlisting}
\end{minipage}

\section{Cross referencing}\label{sec8}

Environments such as figure, table, equation, align can have a label
declared via the \verb+\label{#label}+ command. For figures and table
environments one should use the \verb+\label{}+ command inside or just
below the \verb+\caption{}+ command.  One can then use the
\verb+\ref{#label}+ command to cross-reference them. As an example, consider
the label declared for Figure \ref{fig1} which is
\verb+\label{fig1}+. To cross-reference it, use the command
\verb+ Figure \ref{fig1}+, for which it comes up as
``Figure~\ref{fig1}".

\subsection{Details on reference citations}\label{subsec3}

With standard numerical .bst files, only numerical citations are possible.
With an author-year .bst file, both numerical and author-year citations are possible.

If author-year citations are selected, \verb+\bibitem+ must have one of the following forms:

{\footnotesize%
\begin{verbatim}
\bibitem[Jones et al.(1990)]{key}...
\bibitem[Jones et al.(1990)Jones,
                Baker, and Williams]{key}...
\bibitem[Jones et al., 1990]{key}...
\bibitem[\protect\citeauthoryear{Jones,
                Baker, and Williams}
                {Jones et al.}{1990}]{key}...
\bibitem[\protect\citeauthoryear{Jones et al.}
                {1990}]{key}...
\bibitem[\protect\astroncite{Jones et al.}
                {1990}]{key}...
\bibitem[\protect\citename{Jones et al., }
                1990]{key}...
\harvarditem[Jones et al.]{Jones, Baker, and
                Williams}{1990}{key}...
\end{verbatim}}

This is either to be made up manually, or to be generated by an
appropriate .bst file with BibTeX. Then,

{%
\begin{verbatim}
                    Author-year mode
                        || Numerical mode
\citet{key} ==>>  Jones et al. (1990)
                        || Jones et al. [21]
\citep{key} ==>> (Jones et al., 1990) || [21]
\end{verbatim}}

\noindent
Multiple citations as normal:

{%
\begin{verbatim}
\citep{key1,key2} ==> (Jones et al., 1990;
                         Smith, 1989)||[21,24]
        or (Jones et al., 1990, 1991)||[21,24]
        or (Jones et al., 1990a,b)   ||[21,24]
\end{verbatim}}

\noindent
\verb+\cite{key}+ is the equivalent of \verb+\citet{key}+ in author-year mode
and  of \verb+\citep{key}+ in numerical mode. Full author lists may be forced with
\verb+\citet*+ or \verb+\citep*+, e.g.

{%
\begin{verbatim}
\citep*{key} ==>> (Jones, Baker, and Mark, 1990)
\end{verbatim}}

\noindent
Optional notes as:

{%
\begin{verbatim}
\citep[chap. 2]{key}     ==>>
        (Jones et al., 1990, chap. 2)
\citep[e.g.,][]{key}     ==>>
        (e.g., Jones et al., 1990)
\citep[see][pg. 34]{key} ==>>
        (see Jones et al., 1990, pg. 34)
\end{verbatim}}

\noindent
(Note: in standard LaTeX, only one note is allowed, after the ref.
Here, one note is like the standard, two make pre- and post-notes.)

{%
\begin{verbatim}
\citealt{key}   ==>> Jones et al. 1990
\citealt*{key}  ==>> Jones, Baker, and
                        Williams 1990
\citealp{key}   ==>> Jones et al., 1990
\citealp*{key}  ==>> Jones, Baker, and
                        Williams, 1990
\end{verbatim}}

\noindent
Additional citation possibilities (both author-year and numerical modes)

{%
\begin{verbatim}
\citeauthor{key}       ==>> Jones et al.
\citeauthor*{key}      ==>> Jones, Baker, and
                                Williams
\citeyear{key}         ==>> 1990
\citeyearpar{key}      ==>> (1990)
\citetext{priv. comm.} ==>> (priv. comm.)
\citenum{key}          ==>> 11 [non-superscripted]
\end{verbatim}}

\noindent
Note: full author lists depends on whether the bib style supports them;
if not, the abbreviated list is printed even when full requested.

\noindent
For names like della Robbia at the start of a sentence, use

{%
\begin{verbatim}
\Citet{dRob98}      ==>> Della Robbia (1998)
\Citep{dRob98}      ==>> (Della Robbia, 1998)
\Citeauthor{dRob98} ==>> Della Robbia
\end{verbatim}}

\noindent
The following is an example for \verb+\cite{...}+: \cite{rahman2019centroidb}. Another example for \verb+\citep{...}+: \citep{bahdanau2014neural,imboden2018cardiorespiratory,motiian2017unified,american2013acsm,murphy2012machine,ji20123d}.
Sample cites here \cite{krizhevsky2012imagenet,horvath2018dna} and \cite{pyrkov2018quantitative}, \cite{wang2018face}, \cite{lecun2015deep,zhang2018fine,ravi2016deep}.

\section{Lists}\label{sec9}

List in \LaTeX{} can be of three types: numbered, bulleted and unnumbered. ``enumerate'' environment produces numbered list,
``itemize'' environment produces bulleted list and ``unlist'' environment produces unnumbered list.
In each environments, new entry is added via the \verb+\item+ command.
\begin{enumerate}[1.]
\item This is the 1st item

\item Enumerate creates numbered lists, itemize creates bulleted lists and
unnumerate creates unnumbered lists.
\begin{enumerate}[(a)]
\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.

\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.
\begin{enumerate}[(ii)]
\item Third level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.

\item Third level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.
\end{enumerate}

\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.

\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.
\end{enumerate}

\item Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.

\item Numbered lists continue.
\end{enumerate}
List in \LaTeX{} can be of three types: enumerate, itemize and description.
In each environments, new entry is added via the \verb+\item+ command.
\begin{itemize}
\item First level bulleted list. This is the 1st item

\item First level bulleted list. Itemize creates bulleted lists and description creates unnumbered lists.
\begin{itemize}
\item Second level dashed list. Itemize creates bulleted lists and description creates unnumbered lists.

\item Second level dashed list. Itemize creates bulleted lists and description creates unnumbered lists.

\item Second level dashed list. Itemize creates bulleted lists and description creates unnumbered lists.
\end{itemize}

\item First level bulleted list. Itemize creates bulleted lists and description creates unnumbered lists.

\item First level bulleted list. Bullet lists continue.
\end{itemize}

\noindent
Example for unnumbered list items:

\begin{unlist}
\item Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text.

\item Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text.

\item sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text.
\end{unlist}

\section{Examples for theorem like environments}\label{sec10}

For enunciation like environments, we require \verb+amsthm+ package.

\begin{theorem}[Theorem subhead]\label{thm1}
Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text.
Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text.
Example theorem text.
\end{theorem}

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue.

\begin{proposition}
Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text.
Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text.
\end{proposition}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante.

\begin{example}
Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem.
\end{example}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante.

\begin{remark}
Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem.
\end{remark}

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst.

\begin{definition}[Definition sub head]
Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text.
\end{definition}

Apart from the above styles, we have \verb+\begin{proof}+ \verb+...+ \verb+\end{proof}+ environment - with proof head in italic style and body text in roman font with an open square at the end of each proof environment.

\begin{proof}Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text.
\end{proof}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi.

\begin{proof}[Proof of Theorem~{\upshape\ref{thm1}}]
Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text.
\end{proof}

\noindent
For quote environment, one has to use\newline \verb+\begin{quote}...\end{quote}+
\begin{quote}
Quoted text example. Aliquam porttitor quam a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
\end{quote}
Donec congue. Maecenas urna mi, suscipit in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl (refer Figure~\ref{fig3}). Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Etiam ligula arcu,
elementum a, venenatis quis, sollicitudin sed, metus. Donec nunc pede, tincidunt in, venenatis vitae, faucibus vel (refer Table~\ref{tab3}).

\section{Conclusion}

Some Conclusions here.

%%%%%%%%%%%%%%

\appendix

\section{Section title of first appendix}\label{sec11}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

\subsection{Subsection title of first appendix}\label{subsec4}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

\subsubsection{Subsubsection title of first appendix}\label{subsubsec3}

Example for Unnumbered figure:

\begin{figure}[!h]
\centering
\includegraphics{02}
\end{figure}

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis.

\section{Section title of second appendix}\label{sec12}%

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
\begin{figure}[!b]
\centering
\includegraphics{02}
\caption{This is an example for appendix figure\label{fig4}}
\end{figure}

\subsection{Subsection title of second appendix}\label{subsec5}

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non odio.
Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce sed justo
eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec
pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin et quam.
Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien turpis,
fermentum vel, eleifend faucibus, vehicula eu, lacus.

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non odio.
Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce sed justo
eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec
pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin et quam.
Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien turpis,
fermentum vel, eleifend faucibus, vehicula eu, lacus.

\subsubsection{Subsubsection title of second appendix}\label{subsubsec4}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
vehicula augue eu neque.

\begin{table}[t]%
\caption{This is an example of Appendix table showing food requirements of army, navy and airforce\label{tab4}}%
\begin{tabular*}{\textwidth}{@{\extracolsep\fill}lcc@{}}%
\toprule
col1 head & col2 head & col3 head \\
\midrule
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
\botrule
\end{tabular*}
\end{table}

Example for an equation inside appendix:
\begin{align}
 p &= \frac{\gamma^{2} - (n_{C} -1)H}{(n_{C} - 1) + H - 2\gamma} \label{1eq:hybobo:pfromgH} \\
 \theta &= \frac{(\gamma - H)^{2}(\gamma - n_{C} -1)^{2}}{(n_{C} - 1 + H - 2\gamma)^{2}} \label{2eq:hybobo:tfromgH}\; .
\end{align}

\section{Example of another appendix section}\label{sec13}%

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris
\begin{equation}
\mathcal{L} = i \bar{\psi} \gamma^\mu D_\mu \psi
    - \frac{1}{4} F_{\mu\nu}^a F^{a\mu\nu} - m \bar{\psi} \psi
\label{eq26}
\end{equation}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

%% Example for unnumbered table inside appendix
\begin{table}
\caption{}{%
\begin{tabular}{lcc}%
\toprule
col1 head & col2 head & col3 head \\
\midrule
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
\botrule
\end{tabular}}{}
\end{table}

\section*{Acknowledgments}
The authors thank the anonymous reviewers for their valuable suggestions. This work is supported in part by funds from the National Science Foundation (NSF: \# 1636933 and \# 1920920).

\section*{Funding}
Funding information here.

%\bibliographystyle{plain}
%\bibliography{reference}

\begin{thebibliography}{10}

\bibitem{bahdanau2014neural}
Dzmitry Bahdanau, Kyunghyun Cho, and Yoshua Bengio.
\newblock Neural machine translation by jointly learning to align and
  translate.
\newblock {\em arXiv preprint arXiv:1409.0473}, 2014.

\bibitem{horvath2018dna}
Steve Horvath and Kenneth Raj.
\newblock Dna methylation-based biomarkers and the epigenetic clock theory of
  ageing.
\newblock {\em Nature Reviews Genetics}, 19(6):371, 2018.

\bibitem{imboden2018cardiorespiratory}
Mary~T Imboden, Matthew~P Harber, Mitchell~H Whaley, W~Holmes Finch, Derron~L
  Bishop, and Leonard~A Kaminsky.
\newblock Cardiorespiratory fitness and mortality in healthy men and women.
\newblock {\em Journal of the American College of Cardiology},
  72(19):2283--2292, 2018.

\bibitem{ji20123d}
Shuiwang Ji, Wei Xu, Ming Yang, and Kai Yu.
\newblock 3d convolutional neural networks for human action recognition.
\newblock {\em IEEE Transactions on Pattern Analysis and Machine Intelligence},
  35(1):221--231, 2012.

\bibitem{krizhevsky2012imagenet}
Alex Krizhevsky, Ilya Sutskever, and Geoffrey~E Hinton.
\newblock Imagenet classification with deep convolutional neural networks.
\newblock In {\em Advances in Neural Information Processing Systems}, pages
  1097--1105, 2012.

\bibitem{lecun2015deep}
Yann LeCun, Yoshua Bengio, and Geoffrey Hinton.
\newblock Deep learning.
\newblock {\em Nature}, 521(7553):436, 2015.

\bibitem{motiian2017unified}
Saeid Motiian, Marco Piccirilli, Donald~A Adjeroh, and Gianfranco Doretto.
\newblock Unified deep supervised domain adaptation and generalization.
\newblock In {\em Proceedings of the IEEE International Conference on Computer
  Vision}, pages 5715--5725, 2017.

\bibitem{murphy2012machine}
Kevin~P Murphy.
\newblock {\em Machine learning: A probabilistic perspective}.
\newblock MIT press, 2012.

\bibitem{american2013acsm}
American~College of~Sports~Medicine et~al.
\newblock {\em ACSM's guidelines for exercise testing and prescription}.
\newblock Lippincott Williams \& Wilkins, 2013.

\bibitem{pyrkov2018quantitative}
Timothy~V Pyrkov, Evgeny Getmantsev, Boris Zhurov, Konstantin Avchaciov,
  Mikhail Pyatnitskiy, Leonid Menshikov, Kristina Khodova, Andrei~V Gudkov, and
  Peter~O Fedichev.
\newblock Quantitative characterization of biological age and frailty based on
  locomotor activity records.
\newblock {\em Aging (Albany NY)}, 10(10):2973, 2018.

\bibitem{rahman2019centroidb}
Syed~Ashiqur Rahman and Donald Adjeroh.
\newblock Centroid of age neighborhoods: A generalized approach to estimate
  biological age.
\newblock In {\em 2019 IEEE EMBS International Conference on Biomedical \&
  Health Informatics (BHI)}, pages 1--4. IEEE, 2019.

\bibitem{ravi2016deep}
Daniele Rav{\`\i}, Charence Wong, Fani Deligianni, Melissa Berthelot, Javier
  Andreu-Perez, Benny Lo, and Guang-Zhong Yang.
\newblock Deep learning for health informatics.
\newblock {\em IEEE {J}ournal of {B}iomedical and {H}ealth {I}nformatics},
  21(1):4--21, 2016.

\bibitem{wang2018face}
Zongwei Wang, Xu~Tang, Weixin Luo, and Shenghua Gao.
\newblock Face aging with identity-preserved conditional generative adversarial
  networks.
\newblock In {\em Proceedings of the IEEE Conference on Computer Vision and
  Pattern Recognition}, pages 7939--7947, 2018.

\bibitem{zhang2018fine}
Ke~Zhang, Na~Liu, Xingfang Yuan, Xinyao Guo, Ce~Gao, and Zhenbing Zhao.
\newblock Fine-grained age estimation in the wild with attention {LSTM}
  networks.
\newblock {\em arXiv preprint arXiv:1805.10445}, 2018.

\end{thebibliography}

%USE THE BELOW OPTIONS IN CASE YOU NEED AUTHOR YEAR FORMAT.
%\bibliographystyle{abbrvnat}
%\bibliography{reference}

\end{document}

### 2024-07-24 tex/title_page.tex
after resubmission

\documentclass[numbers,webpdf,imaman]{ima-authoring-template}%

\usepackage{latexsym}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{mathptmx}
\usepackage{placeins}
% \usepackage{fullpage}
\usepackage[left=16mm, right=16mm, top=23mm, bottom=23mm]{geometry}
%
%****************************************************************************
% AUTHOR: You may want to use some of these packages. (Optional)
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{epsfig}

\usepackage{caption}
\usepackage{standalone}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{calc}
\usetikzlibrary{shapes}

% \usepackage[round]{natbib}
% \usepackage[linesnumbered,lined,ruled,commentsnumbered]{algorithm2e}

\newcommand{\specialcellr}[2][r]{%
  \begin{tabular}[#1]{@{}r@{}}#2\end{tabular}}

%****************************************************************************

%
%****************************************************************************
% AUTHOR: If you do not wish to use hyperlinks, then just comment
% out the hyperref usepackage commands below.

%% This version of the command is used if you use pdflatex. In this case you
%% cannot use ps or eps files for graphics, but pdf, jpeg, png etc are fine.

% \usepackage[
%   colorlinks=true,
%   urlcolor=blue,
%   citecolor=black,
%   anchorcolor=black,
%   linkcolor=black,
% ]{hyperref}
% \usepackage{xurl}

%% The next versions of the hyperref command are used if you adopt the
%% outdated latex-dvips-ps2pdf route in generating your pdf file. In
%% this case you can use ps or eps files for graphics, but not pdf, jpeg, png etc.
%% However, the final pdf file should embed all fonts required which means that you have to use file
%% formats which can embed fonts. Please note that the final PDF file will not be generated on your computer!

%%\usepackage[dvips,colorlinks=true,urlcolor=blue,citecolor=black,%
%% anchorcolor=black,linkcolor=black]{hyperref}
%****************************************************************************

%
%****************************************************************************
%*
%* AUTHOR: YOUR CALL!  Document-specific macros can come here.
%*
\setcitestyle{authoryear,open={(},close={)}} %Citation-related commands
%****************************************************************************
% Packages added manually
\usepackage{xcolor}
\usepackage{multirow}
%#########################################################
%*
%*  The Document.
%*
\begin{document}

\DOI{DOI HERE}
\copyrightyear{2023}
\vol{00}
\pubyear{2024}
\access{Advance Access Publication Date: Day Month Year}
\appnotes{Paper}
\copyrightstatement{Published by Oxford University Press on behalf of the Institute of Mathematics and its Applications. All rights reserved.}
\firstpage{1}

\title[Optimising Heterogeneous Ambulance Fleet Allocations in Jakarta]{Optimising Heterogeneous Ambulance Fleet Allocations in Jakarta}

\author{
Geraint Ian Palmer*, Mark Tuson, Vincent Knight, Paul Harper, and Sarie Brice
\address{
    \orgdiv{School of Mathematics},
    \orgname{Cardiff University},
    \orgaddress{\street{Abacws},
    \postcode{CF24 4AG},
    \state{Cardiff, Wales},
    \country{United Kingdom}
}}}
\author{Leanne Smith
\address{
    \orgname{Welsh Ambulance Services NHS Trust},
    \orgaddress{\street{Beacon House},
    \postcode{NP44 3AB},
    \state{Cwmbran, Wales},
    \country{United Kingdom}
}}}
\author{Daniel Gartner
\address{
    \orgname{Aneurin Bevan University Health Board},
    \orgaddress{\street{Lodge Road},
    \postcode{NP18 3XQ},
    \state{Caerleon, Wales},
    \country{United Kingdom}
}}}

\corresp[*]{Corresponding author: \href{email:palmergi1@cardiff.ac.uk}{palmergi1@cardiff.ac.uk}}

\received{Date}{0}{Year}
\revised{Date}{0}{Year}
\accepted{Date}{0}{Year}

\abstract{Ambulance services have a duty of care to the clinical outcomes of the
population they serve, and therefore aim to maximise the chances of survival and
improve patient outcomes following a medical emergency. Despite this, although
the ambulance allocation problem has been widely studied, it has predominantly
focused on minimising response times or maximising coverage alone, and not
explicitly for considering patient outcomes. In this paper we propose a
modelling approach to consider where to best allocate different types of
emergency response vehicles in order to maximise patient outcomes within a
heterogeneous population. To achieve this, we develop a metaheuristic algorithm
for finding better fleet allocations which is used in conjunction with a
discrete-event simulation model of ambulance services with heterogeneous
vehicles. A major contribution of this metaheuristic is the numerical solution
of a system of equations to approximate the utilisation of vehicles.
Traditionally this utilisation is problematic as it is both an input and an
output of the allocation of vehicles. Our approach is informed by, and tested
on, real-world data from Jakarta, Indonesia. Using our developed models,
decision makers are better able to understand ambulance fleet capacity needs and
allocations, and their impact on patient outcomes.}

\keywords{Ambulance Planning; Emergency Medical Services; Simulation;
Health Care; Optimisation; Mathematical Modelling}

\maketitle

\end{document}

### 2024-07-26 response/response.tex
addresses a number of comments from reviewer 1

\documentclass{article}

\usepackage{fullpage}
\usepackage{parskip}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{hyperref}

\title{Responses to Reviewers}
\author{}
\date{}

\begin{document}

\maketitle

We would like to express our gratitude to the editor, associate editor, and reviewers for facilitating the review process, and for their useful suggestions and remarks on the manuscript.

This document will address the specific concerns raised.

\section{Responses to the Editor}

\begin{itemize}

\item The editor wrote:
\begin{quote}
``Please cite relevant, recent work published in this journal. This is by no means mandatory, but it will help to position your work in the context of research published in the journal.''
\end{quote}

\item The editor wrote:
\begin{quote}
``The manuscript does not sufficiently discuss the managerial implications of the work. Please develop this further in your revision.''
\end{quote}

\end{itemize}

\section{Responses to the Associate Editor}
There were no comments by the associate editor. We would like to thank the associate editor for their work facilitating the review process.

\section{Responses to the Reviewers}

\subsection{Comments from Reviewer 1}

\begin{itemize}

\item Reviewer 1 wrote:
\begin{quote}
``No Verification \& Validation of the simulation model is described. At least some discussion regarding the validity of the model has to be included in the case study.''
\end{quote}

\item Reviewer 1 wrote:
\begin{quote}
``No measures of variance are given for the simulation results in Tables 1 and 4. Especially since repeated simulation results were conducted, measures of variance should be included to investigate the statistical significance of the results.''
\end{quote}

\item Reviewer 1 wrote:
\begin{quote}
``Algorithm 1: Updating the mutation rate m by subtracting the maximum of m0 and a second number from m0 will lead to a negative (or zero) mutation rate, which does not make sense. Probably should be min instead of max.''
\end{quote}
The algorithm is correct. Line 5 of the algorithm states that $m_0 \leftarrow \lceil m_0 - \max(m_0, i \times c) \rceil$. Without the ceiling function, this would have the effect of linearly decreasing wiht $i$ until 0, and then remaining at 0 thereafter. With the ceiling function, the linear part becomes a stepwise decent to 0, and then 0 thereafter.

\item Reviewer 1 wrote:
\begin{quote}
``Unclear, why the subchapter 4.1 on survival functions is necessary, since traditional coverage targets are used in the case study and the model formulation does not depend on characteristics of the survival functions. While survival functions are an important performance metric in the reviewers opinion, it is not clear why the reader needs this information to follow the remainder of the paper.''
\end{quote}
We disagree with the reviewer here. In the case study, traditional targets such a mean response times are reported alongside survival measures, based on the survival functions. However, a key part of the paper is the optimisation process, and all scenarios compared are `optimised' scenarios that result from this process. The objective function of the optimisation process maximises survival, based on survival functions. This is evnidenced in Equations 6 and 7, which are component parts of the objective function given in Equation 5. Therefore the survival functions are central to the work presented.

\item Reviewer 1 wrote:
\begin{quote}
``p. 8/9: A short explanation, of what the MINPACK hybrd and hybrj algorithms do and how they work should be included either directly or in the appendix.''
\end{quote}

\item Reviewer 1 wrote:
\begin{quote}
``There is no discussion of how well the proposed model estimates survival probabilities. This could be easily done by calculating the survival probabilites based on the simulated response times and comparing the simulated survival probabilities with the ones calculated by the MESLMHPHF. It is also not clear, why e.g. in Table 4 the expected Survival from the MESLMHPHF was used instead of the more realistic survival based on the simulation results.''
\end{quote}

\item Reviewer 1 wrote:
\begin{quote}
``Formulating the MESLMHPHF as an explicit optimization model with an objective function and constraints would make it easier to follow the model logic than it is with the separation of the formulation over multiple pages.''
\end{quote}
This has now been given as an Appendix, while the body of the paper presents the model in a way that explains and justifies each step.

\item Reviewer 1 wrote:
\begin{quote}
``p.1 l. 37 ‘driven by inadequate a lack of financial investment and human capital’ (delete inadequate)''
\end{quote}
Fixed.

\item Reviewer 1 wrote:
\begin{quote}
``p.5, l. 21: $J_{yk}$ was not defined, only $J_k$''
\end{quote}
Fixed, all instances should the $J_k$.

\item Reviewer 1 wrote:
\begin{quote}
``Fig. 10: Both plots should have the same y-axis interval and ideally start
at 0''
\end{quote}
The scale of Fig 10.b is so small that the lines would not be able to be differentiated if it had the same scale as Fig 10.a; and would be impossible to see at all if the y-axis starts at 0. Instead, we have included a sentence in the caption to note that the y-axes are different for both plots.

\item Reviewer 1 wrote:
\begin{quote}
``Fig. 11: A1 survival does not increase monotonically, especially in the single vehicle type scenario. A discussion of possible causes for this counterintuitive result should be included.''
\end{quote}
Thank you for this comment. We have now added some sentences explaining this phenomenon, calling back to the behaviour of the survival function itself and arguing that this is an upper bound for the survival of A1 patients.

\item Reviewer 1 wrote:
\begin{quote}
``Equations 8 and 9: It is stated, that $\beta_{p a_1 a_2}$ indicates, that a vehicle from a1 can reach p quicker, than a vehicle from a2. The equations 8 and 9 however use $\leq$ and not < , which is not consistent with the description.''
\end{quote}
Thank you for spotting this. As the $\beta_{p a_1 a_2}$ define an ordering of the ambulance locations, it does not matter whether they are defined by $\leq$ or <, as we assume that no two distinct ambulance locations are equidistant from a pickup location. We have now added this point to the text.

\item Reviewer 1 wrote:
\begin{quote}
``p. 8, L.25-27: The reason why the fact that the preferences depending on the travel times should reduce the impact of including location specific service rates is not obvious and shuould be expanded on.''
\end{quote}
We have now added a sentence explaining this.

\item Reviewer 1 wrote:
\begin{quote}
``Algorithm 1: Even though it is obvious that the solutions are very likely ranked by decreasing expected survival probability it should be stated clearly, how the population is ranked according to Equation 5.''
\end{quote}
Fixed.

\item Reviewer 1 wrote:
\begin{quote}
``Table 4: The optimised solution shows (slightly) higher mean response times for scenarios D13 and D19. This should be discussed.''
\end{quote}
This is discussed directly in the paragraph discussing the table: ``It is worth noting that for the simulation based KPIs (mean response times and percent abandoned), there is not too much difference between the currently used allocation and the better allocation, showing that these KPIs may not be good approximations of survival.''

\item Reviewer 1 wrote:
\begin{quote}
``The relaxation of the assumption of location independent service rates and and time-independent arrivale rates in the MESLMHPHF could be remarked upon in the outlook''
\end{quote}
This has now been added to the discussion section.

\item Reviewer 1 wrote:
\begin{quote}
``Reframing subchapter 6.3 as e.g. demand scenario definition would help the reading flow''
\end{quote}

\end{itemize}

\subsection{Comments from Reviewer 2}

\begin{itemize}

\item Reviewer 2 wrote:
\begin{quote}
``The introduction makes a case for using survival vs reponse time as a measure for ambulance allocation (although this is not the only measure in the results). What would this do to fairness in the allocations? Would far away patients with serious conditions (ie. Needing a fast response time) just be discarded as it is expensive to cater for them (in terms of overall survival of the population)?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``Is the survival function the same for $K_a$ patients regardless of which vehicle visits them? What is the rationale for this? It seems that primary vehicles would be better prepared to deal with complications, etc.''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``Have you given any consideration to using an exact method to solve the optimisation problem solved in Algorithm 1? Is there any alternative in the literature?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P11 L43 – What is the rationale of using a triangular distribution for travel times? In general, what is the basis of all the travel time assumptions presented in this section?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P12 5.4 – For patients that might not need to travel to a hospital, would they still need a primary vehicle after a secondary vehicle tends to them? If not the logic would not be separate entirely.''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``It seems there are two conflicting methods here, the optmisiation with the EA, which uses some averages, etc. And a more comprehensive DES simulation; however the latter is not used for the optimisation. Did you consider simulation optimisation or simheuristics to address this?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P15 Paragraph starting L36. Can you give more detail on the modelling of the call arrival rates? For rare specialties, and given the low usage of ambulances in Indonesia, you might not have enough data to give an accurate representation of actual demand. Especially, for future demand where visibility is increased, how will this change for example among richer/poorer parts of the city? I acknowledge it is impossible to include everything in your experiments, but how do you expect these limitations might affect your results?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``Could you conduct any validation of your results with real data from eg. Ambulance response times with the baseline allocation?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P4 L34 – Note that IF .... it is possible''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P4 L41-42 – Travel back to ITS (not it’s)''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P7 L29 – equations in APPENDIX A.''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P8 Fig 3 – Include y axis scale''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P8 L27 – change ending to – This might not have a strong impact on the results of this study.''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P8 L49 – In addition to the package used, can you more explicitly state what optimisation problem was solved here, and by which method?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P12 L26,27 Rewrite, possible a comma is missing there, also L32.''
\end{quote}

\end{itemize}

\end{document}

### 2024-07-26 tex/main.tex
addresses a number of comments from reviewer 1

\documentclass[numbers,webpdf,imaman]{ima-authoring-template}%

\usepackage{soul}
\usepackage{latexsym}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{mathptmx}
\usepackage{placeins}
% \usepackage{fullpage}
\usepackage[left=16mm, right=16mm, top=23mm, bottom=23mm]{geometry}
%
%****************************************************************************
% AUTHOR: You may want to use some of these packages. (Optional)
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{epsfig}

\usepackage{caption}
\usepackage{standalone}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{calc}
\usetikzlibrary{shapes}

% \usepackage[round]{natbib}
% \usepackage[linesnumbered,lined,ruled,commentsnumbered]{algorithm2e}

\newcommand{\specialcellr}[2][r]{%
  \begin{tabular}[#1]{@{}r@{}}#2\end{tabular}}

%****************************************************************************

%
%****************************************************************************
% AUTHOR: If you do not wish to use hyperlinks, then just comment
% out the hyperref usepackage commands below.

%% This version of the command is used if you use pdflatex. In this case you
%% cannot use ps or eps files for graphics, but pdf, jpeg, png etc are fine.

% \usepackage[
%   colorlinks=true,
%   urlcolor=blue,
%   citecolor=black,
%   anchorcolor=black,
%   linkcolor=black,
% ]{hyperref}
% \usepackage{xurl}

%% The next versions of the hyperref command are used if you adopt the
%% outdated latex-dvips-ps2pdf route in generating your pdf file. In
%% this case you can use ps or eps files for graphics, but not pdf, jpeg, png etc.
%% However, the final pdf file should embed all fonts required which means that you have to use file
%% formats which can embed fonts. Please note that the final PDF file will not be generated on your computer!

%%\usepackage[dvips,colorlinks=true,urlcolor=blue,citecolor=black,%
%% anchorcolor=black,linkcolor=black]{hyperref}
%****************************************************************************

%
%****************************************************************************
%*
%* AUTHOR: YOUR CALL!  Document-specific macros can come here.
%*
\setcitestyle{authoryear,open={(},close={)}} %Citation-related commands
%****************************************************************************
% Packages added manually
\usepackage{xcolor}
\usepackage{multirow}
%#########################################################
%*
%*  The Document.
%*
\begin{document}

\DOI{DOI HERE}
\copyrightyear{2023}
\vol{00}
\pubyear{2024}
\access{Advance Access Publication Date: Day Month Year}
\appnotes{Paper}
\copyrightstatement{Published by Oxford University Press on behalf of the Institute of Mathematics and its Applications. All rights reserved.}
\firstpage{1}

\title[Optimising Heterogeneous Ambulance Fleet Allocations in Jakarta]{Optimising Heterogeneous Ambulance Fleet Allocations in Jakarta}

% \author{
% Geraint Ian Palmer*, Mark Tuson, Vincent Knight, Paul Harper, and Sarie Brice
% \address{
%     \orgdiv{School of Mathematics},
%     \orgname{Cardiff University},
%     \orgaddress{\street{Abacws},
%     \postcode{CF24 4AG},
%     \state{Cardiff, Wales},
%     \country{United Kingdom}
% }}}
% \author{Leanne Smith
% \address{
%     \orgname{Welsh Ambulance Services NHS Trust},
%     \orgaddress{\street{Beacon House},
%     \postcode{NP44 3AB},
%     \state{Cwmbran, Wales},
%     \country{United Kingdom}
% }}}
% \author{Daniel Gartner
% \address{
%     \orgname{Aneurin Bevan University Health Board},
%     \orgaddress{\street{Lodge Road},
%     \postcode{NP18 3XQ},
%     \state{Caerleon, Wales},
%     \country{United Kingdom}
% }}}

% \corresp[*]{Corresponding author: \href{email:palmergi1@cardiff.ac.uk}{palmergi1@cardiff.ac.uk}}

\received{Date}{0}{Year}
\revised{Date}{0}{Year}
\accepted{Date}{0}{Year}

\abstract{Ambulance services have a duty of care to the clinical outcomes of the
population they serve, and therefore aim to maximise the chances of survival and
improve patient outcomes following a medical emergency. Despite this, although
the ambulance allocation problem has been widely studied, it has predominantly
focused on minimising response times or maximising coverage alone, and not
explicitly for considering patient outcomes. In this paper we propose a
modelling approach to consider where to best allocate different types of
emergency response vehicles in order to maximise patient outcomes within a
heterogeneous population. To achieve this, we develop a metaheuristic algorithm
for finding better fleet allocations which is used in conjunction with a
discrete-event simulation model of ambulance services with heterogeneous
vehicles. A major contribution of this metaheuristic is the numerical solution
of a system of equations to approximate the utilisation of vehicles.
Traditionally this utilisation is problematic as it is both an input and an
output of the allocation of vehicles. Our approach is informed by, and tested
on, real-world data from Jakarta, Indonesia. Using our developed models,
decision makers are better able to understand ambulance fleet capacity needs and
allocations, and their impact on patient outcomes.}

\keywords{Ambulance Planning; Emergency Medical Services; Simulation;
Health Care; Optimisation; Mathematical Modelling}

\maketitle

\section{Introduction and Background}\label{sec:intro}
Time-critical conditions (TCCs) are a substantial cause of mortality
worldwide, responsible for an estimated 54\% of all deaths \citep{FraserBMJ}.
Emergency medical services (EMS) play a pivotal role in responding to TCCs and
saving patients from life-threatening medical conditions. Most EMS research,
especially  planning to meet increasing ambulance demands, tends to be
focussed on high-income countries including, for example, Australia
\citep{lowthian2011increasing}, Germany \citep{veser2015demographic}, and USA
\citep{birmingham2021trends}. In contrast, many low and middle-income countries
(LMICs) lack an organised EMS system, with most ambulances used purely for
patient transport and not as an emergency care vehicle \citep{plummer2017ems}.
However, the burden of deaths due to TCCs is much greater in LMICs than in
high-income countries, the difference being around threefold \citep{ChangPMC}.
For traffic accidents, strokes and heart attacks, rapid access to appropriate
treatment is especially vital.

The application of our research is focused on Indonesia, supported by an
award for Global Challenges Research Funds (GCRF). In Indonesia, an LMIC,
there has been very little prior work on developing an EMS strategy, driven
by \textcolor{red}{\st{inadequate}} a lack of financial investment and human capital
\citep{plummer2017ems,pusponegoro2003terrorism,yusvirazi2018state}.
Related research by the authors \citep{BriceSyaribahNoor2022Esui} has
highlighted particular challenges in Indonesia as a lack of a single
coordinated EMS, the ability or willingness to pay for an ambulance, a large
geographical area, and areas of severe traffic congestion especially in
Jakarta, the capital.

When our research programme commenced in October 2019, ambulance services in
Jakarta were provided by many disparate, mostly private providers that charge
patients for their use. We initially partnered with Ambulans 118, a
non-government charitable ambulance service established in 2005 by the
Indonesian Surgeons’ Association, that  currently operates in five cities across
Indonesia: Jakarta, Palembang, Yogyakarta, Surabaya and Makasser. Unlike private
providers, Ambulans 118 suggests that a donation is made for use of its EMS
vehicles, but otherwise provides free emergency medical care, it is also leading
on paramedic training across the country.

The overarching goal of our research was to work with the Indonesian Government,
the different ambulance providers, and hospitals, to help them forecast
emergency demand and make critical decisions on the best types, capacities and
geographical allocations of emergency vehicles within a potentially co-ordinated
EMS system, starting with Jakarta. To facilitate this, in collaboration with
Ambulans 118 we organised workshops in Jakarta in February 2020, which were
attended by over 170 people including doctors, nurses, paramedics, academics,
and officials from the Indonesian Ministry of Health. This paper presents some
aspects of the overall research programme, specifically focusing on the
development of a simulation model and allocation metaheuristic for maximising
patient survival.

Fundamental to the project and the research described in this paper was an
initial study that we carried out in order to better understand patient needs
and the barriers to use of ambulances in Indonesia. Throughout the month of
December 2020, we undertook comprehensive surveys in Emergency Departments
(EDs) across Jakarta and published the first known study in to EMS demand
within the country \citep{BriceSyaribahNoor2022Esui}. Our study showed that the
utilisation of ambulances by patients attending EDs is very low and of
concern. The low utilisation is contributed by patients' lack of awareness of
available ambulance services, patients' disinclination to use ambulances due to
high costs, and long response times.
All of these barriers impact on patient outcomes, especially for those with
life-threatening conditions. For example, more trauma patients took
a car-share ride (20\%) or motorcycle (20\%) to reach the ED than an ambulance
(just 10\%), while only 14\% of critical cardiovascular patients used an
ambulance compared to 67\% travelling to hospital by private car or a
car-share ride.

The first contribution of this paper is an extension the work of
\citet{Knight2012918} to consider both heterogeneous patients and heterogeneous
fleets. Here a maximal survival objective is formulated, from vehicle
allocations and utilisations, which are themselves found by numerically
solving a relationship on the vehicles' share of the demand. This is then used
in a metaheuristic to find improved allocations that aim to maximise patient
survival.
Secondly, we develop a comprehensive sequential discrete event EMS
simulation that models and evaluates heterogeneous fleet allocations of
emergency vehicles by utilising a novel approach in which transit jobs, rather
than patients, are framed as the queueing `customers'. Our methodology
utilises two discrete event simulations of the same system that are run
sequentially and together combine to form the logic of a single simulation of
a heterogeneous fleet; and key performance indicators (KPIs) are calculated by
making use of survival functions.
Thirdly, we demonstrate the use and impact of our modelling framework applied
to current and proposed ambulance allocations in the city of Jakarta, thus
supporting Government-level decision making with an overall goal to improve
the lives of those living in LMICs. This enables decision support and
managerial insights in Indonesia, although the developed modelling framework
could be readily applied to other locations.

The paper is structured as follows:
Section~\ref{sec:litreview} gives an overview of the literature on modelling
ambulance allocations.
Section~\ref{sec:problem_description} describes the current ambulance service
behaviour and sets out the mathematical notation used throughout the paper.
Section~\ref{sec:optimising} describes an
optimisation model, including the survival objective, utilisation
considerations, and a metaheuristic algorithm to find better fleet allocations.
Section~\ref{sec:simulation} outlines the logic of the sequential discrete
event simulation models for heterogeneous fleets.
Section~\ref{sec:jakarta} gives a case study, describing the current emergency
service situation in Jakarta and applying the optimisation and simulation
results.
Section~\ref{sec:discussion} discusses the findings and contributions of the
paper.

\section{Literature Review}\label{sec:litreview}
There is a rich literature on operational research applied to EMS location and
related allocation problems. Literature reviews that include these topics are
\citet{aringhieri2017emergency, belanger2019recent, farahani2019or, Li2011, Liu2021, reuter2017logistics, mukhopadhyay22}
and \citet{wang2021emergency}. Ambulance allocation problems such as that
concerned in this paper can be categorised as a \textit{preparedness}
problem \citep{mukhopadhyay22}.

Emergency medical services can be evaluated using different performance
measures with coverage and response time being the predominant metrics
\citep{Mclay2010evaluating}. However, missing a response time target by just
one second, for example, would be considered as a `failure' in many models
with no appreciation at all of the impact on patient survival or outcome
\citep{Mclay2010evaluating}. This seems somewhat restrictive and short-sighted,
which is why we focus on survival as a metric and
compare and contrast papers using that metric with our approach.
Here we consider some closely related papers, providing a summary of each and
comparing and contrasting them with our modelling and solution approach.

\citet{amorim2019integrated} propose an integrated strategic and tactical
planning approach which features an optimisation model and a local search
heuristic based on Gaussian Processes. Their methodology is applied to the
city of Porto while reporting on performance metrics such as survival. Our
approach is different because we use a discrete-event simulation framework
to evaluate results from a population based heuristic to determine the number of
ambulances required in each station.

\citet{boutilier2022drone} develop an integrated location-queueing model that
incorporates existing EMS response times in a drone network. They use a
p-median approach and an Erlang loss model. Although survival is not their
main optimisation criterion,  they report how many patients would survive
using their approach.

\citet{Erkut200842} can be considered the first modelling approach
that considers decaying survival probabilities during response time. The
rationale is, because survival can be thought of as a more robust and generic
objective for EMS performance measurement than coverage or average response
time. In an application of a recently developed Maximal Survival Location
Problem model (MSLP), the authors use data from Edmonton, Canada and show that
maximising survival is superior to other objectives in clinical outcome for
cardiac arrest patients.

\citet{MCormack2015} focus on the optimisation of
EMS vehicle fleet allocation and base station location through the use of a
genetic algorithm (GA) with an integrated EMS simulation model. Their
objective is maximization of the overall expected survival probability across
patient classes. Applications of the model were undertaken using real call
data from the London Ambulance Service. The difference between their modelling
approach and our is that we have a different survival function, focus on a
different heuristic optimisation and evaluate our approach with data from a
developing country, and have higher traffic volume and congestion in our data.

\citet{belanger20} combine optimisation and simulation to find
both ambulance locations and dispatch policies. Here, binary integer programming 
is used to minimise total response time, where ambulance availability, or
utilisation is unknown. In this work this is overcome by iteratively feeding the
solution of the optimisation programme into a simulation to find the
availability parameters, until convergence. Our work differs as we solve for
ambulance utilisation explicitly.

\citet{torodiaz13, torodiaz15} build
optimisation models for both the allocation and the dispatch, or preference,
rules for EMS services. They utilise the hypercube queueing model for ambulance
availability with a integer programming and metaheuristic algorithms. In
particular, they consider a number of different objectives including mean
response times, expected coverage \citep{daskin83}, and the Gini index on
individual response times as measures of fairness in the system.

\citet{Knight2012918} provide a supporting extension
of the MSLP that allows it to be applied more generally to real-world EMS
systems, acknowledging that different patients have varying levels of expected
survival probabilities.  The Maximum Expected Survival Location Model for
Heterogeneous Patients, MESLMHP, allows multiple classes of patient groups to
be defined, where previously only cardiac arrest patients formed part of the
objectives for primary response. In reality any emergency
patient has a necessity for swift attendance, and a timely response to any
incident type may impact on clinical outcome in some way. It is for this
reason MESLMHP is designed to be generic enough to accommodate any number of
patient groups, with each class weighted dependent upon the relative urgency
of the incident.  More than just a contribution to the model’s demand input,
these patients are included in the optimisation when maximising total
population survival probability.
Our work directly extends this work, to include heterogeneous vehicle fleets,
while developing a different fixed-point numerical solution to the problem of
approximating vehicle utilisations.

\section{Problem Statement \& Notation}\label{sec:problem_description}
We first present some notation. We have the following sets:

\begin{itemize}
  \item $\mathcal{P}$ is the set of pick-up locations, indexed by $p$;
  \item $\mathcal{A}$ is the set of ambulance locations, indexed by $a$;
  \item $\mathcal{Y}$ is the set of hospitals, indexed by $y$;
  \item $\mathcal{K}$ is the set of medical specialities, indexed by $k$;
\end{itemize}

\noindent
to incorporate heterogeneous fleets the set of specialities $\mathcal{K}$ is
partitioned into two sets, $\mathcal{K}_A$, those patients that can be seen by
secondary vehicles, and $\mathcal{K}_B$, those patients who are seen by
primary vehicles only.

An allocation is given by the tuple $\left(Z_a, \tilde{Z}_a\right)$, where $Z_a$
is the number of primary vehicles allocated to ambulance location $a$; and
$\tilde{Z}_a$ is the number of secondary vehicles allocated to ambulance
location $a$. Also let:

\begin{itemize}
  \item $\lambda_{pk}$ be the rate at which patients of speciality $k$ make
        calls from pick-up location $p$;
  \item $s_k(t)$ be the survival function function associated with patients of
        speciality $k$;
  \item $\pi_a$ be the average utilisation of primary vehicles stationed at
        location $a$;
  \item $b_{pa}$ be the expected travel time from ambulance location $a$ to
        pick-up location $p$; and let $B_{pa}$ be a random variable representing
        the time it takes for an ambulance to drive this distance;
  \item $C_{py}$ be a random variable representing the travel time from pick-up
        location $p$ to hospital $y$;
  \item $D_{ya}$ be a random variable representing the travel time from hospital
        $y$ to ambulance location $a$;
  \item $F_{pa}$ be a random variable representing the travel time from pick-up
        location $p$ to ambulance location $a$;
  \item $G_k$ be a random variable representing the time the ambulance spends
        with patients of speciality $k$ at the pick-up location;
  \item $J_k$ be a random variable representing the time the ambulance spends
        with patients of speciality $k$ at the hospital;
  \item $\Theta$ be a random variable representing the time the ambulance spends
        re-fuelling, re-stocking, and resting between transit jobs;
  \item $q_{pky}$ be the probability that a patient of speciality $k$ from
        pick-up location $p$ is taken to hospital $y$.
        Note that $\sum_{y \in \mathcal{Y}} q_{pky} < 1$ is possible, that is a
        patient may not go to any hospital, in which case the ambulance returns
        to their ambulance location.
\end{itemize}

To incorporate heterogeneous vehicles, let also $\tilde{\pi}_a$ denote the
average utilisation of secondary vehicles stationed at location $a$; and
$\tilde{b}_{pa}$, $\tilde{B}_{pa}$, $\tilde{C}_{py}$, $\tilde{D}_{ya}$, and
$\tilde{F}_{pa}$ denote the corresponding travel times for secondary vehicles.

Patients will call an ambulance from one of the pick-up locations
$p \in \mathcal{P}$. If \textit{all} ambulances are busy, then that call will be
abandoned and it is assumed that the patient will find their own way to the
hospital through private or public transportation, which is an appropriate and
justified assumption for the case of Jakarta \citep{BriceSyaribahNoor2022Esui}.
Otherwise, a central control centre will dispatch the closest available primary
vehicle. That vehicle will travel from its location to the patient ($B_{pa}$),
spend some time on the scene treating the patient ($G_k$), travel to the
hospital ($C_{py}$), spend some time handing over the patient at the hospital
($J_k$), travel back to it's original location ($D_{ya}$), then spend time
refilling and refuelling ($\Theta$).
There is also a probability that the patient does not need a hospital, and so
the vehicle will travel from the pick-up location back to it's original vehicle
location ($F_{pa}$). These routes are shown visually in
Figure~\ref{fig:travel_routes}. Note that these routes cannot be interrupted,
that is, an ambulance must return to it's station for refilling before it is
available to respond to another call, to ensure it is cleaned, and fully stocked
for it's next call. Note also that $B_{pa}$ and $F_{pa}$ need not necessarily be
equal, as travel times need not necessarily be symmetrical, this could be due to
a number of reasons, such as non-symmetric road networks, and differences in
driver urgency on each leg of the journey.

\begin{figure}
    \centering
    \includestandalone[width=0.475\textwidth]{img/diagram}
    \caption{Ambulance routes for a given transit job.}
    \label{fig:travel_routes}
\end{figure}

Therefore the total time an ambulance is busy will be $T_{paky}$ given by
Equation~\ref{eqn:service_time_withhosp} if a hospital is required, and
service time $T_{pak}$ given by Equation~\ref{eqn:service_time_nohosp} if no
hospital is required.

\begin{align}
T_{paky} &= B_{pa} + G_k + C_{py} + J_k + D_{ya} + \Theta \label{eqn:service_time_withhosp} \\
T_{pak} &= B_{pa} + G_k + F_{pa} + \Theta \label{eqn:service_time_nohosp}
\end{align}

There are two types of vehicle: primary vehicles, typically ambulances, that
must be dispatched to all patients; and secondary vehicles, typically rapid
response motorcycles, that can travel faster than primary vehicles, and are
dispatched to some patients to respond to an emergency faster than the primary
vehicle. Secondary vehicles cannot transport patients, but are used in
conjunction with primary vehicles to reduce response times.
From a resource cost perspective, one primary vehicle costs the same to purchase
and to run as three secondary vehicles.

\section{Optimising Allocations for Maximal Survival}\label{sec:optimising}
Here we outline a method of finding vehicle allocations that maximise the
expected survival of patients.
The problem is, for a given set of vehicle locations (that is ambulance
stations), a given total number of primary vehicles, and a given total number
of secondary vehicles, how many vehicles of each type should be stationed at
each location to maximise expected survival across all patients.

\subsection{Survival Functions}\label{sec:survival}
A key concept here is the survival curve of a patient.
In reality, some emergency incidents do not result in a substantive
deterioration of a patient’s status over time; however, in all situations,
there is a reasonable cut-off beyond which a patient should not expect to have
to wait for care, and mixture of theoretical survival functions and step
functions can be utilised.
Survival probabilities for critical incidents, calculated from a theoretical
monotonically decaying survival function reported in the literature
\citep{Valenzuela20001206} are used to demonstrate an attainable level of
success from a response. One particular survival curve $s(t)$ of
Equation~\ref{eqn:survival_curve}, represents survival until hospital
discharge following cardiac arrest; its origins are explained in detail by
\citet{Knight2012918}, and gives the probability of survival if seen within a
time $t$ in the form of a logistic function. Figure~\ref{fig:survivalfunction}
shows the difference between using this survival curve and a hard cut-off of 8
minutes.
However, hard cut-off curves like Equation~\ref{eqn:survival_cutoff}, with a
cutoff of $L$ can still be used to represent meeting artificially selected
targets, for example for transportation jobs.

\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{img/Survival_Function.pdf}
\caption{Survival function $s(t)$, estimated by \citet{Valenzuela20001206}
compared with $s_L(t)$ with a hard cutoff of $L=8$, represented as a step
function.}
\label{fig:survivalfunction}
\end{figure}

\begin{equation}\label{eqn:survival_curve}
    s(t) = \left(1 + e^{0.26+0.139t}\right)^{-1}
\end{equation}

\begin{equation}\label{eqn:survival_cutoff}
    s_L(t) = \begin{cases}
    1 & \text{ if } 0\leq t \leq L \\
    0 & \text{ if } t > L 
    \end{cases}
\end{equation}

\subsection{Maximal Expected Survival Location Model with Heterogeneous
            Patients and Heterogeneous Fleet}\label{sec:meslmhphf}
Here we propose a Maximal Expected Survival Location Model with Heterogeneous
Patients and Heterogeneous Fleet (MESLMHPHF). It is an extension of the
MESLMHP model given in \citet{Knight2012918}, which did not consider
heterogeneous fleets. This is a model of survival that can be used as an
objective function for optimisation algorithms. It is a weighted expected
survival function as the objective function which considers heterogeneous
patients (different specialities) and heterogeneous fleets (both primary
vehicles, EAs, and secondary vehicles, RRVs), and is constructed by
appropriately summing these survival curves, described in
Section~\ref{sec:survival}, across the population, multiplying by ambulance
availability where appropriate.
\textcolor{red}{The full MESLMHPHF optimisation model is given in
Appendix~\ref{apx:fullmodel}, and its components are described fully here.}
The MESLMHPHF is given by Equation~\ref{eqn:meslmhphf}.

\begin{equation}\label{eqn:meslmhphf}
g\left(Z_a, \tilde{Z}_a\right) =
\sum_{p \in \mathcal{P}} \sum_{a \in \mathcal{A}}
\left( \sum_{k \in \mathcal{K}_A}  w_k \lambda_{pk} \hat{\Psi}_{kpa} +
\sum_{k \in \mathcal{K}_B}  w_k \lambda_{pk} \Psi_{kpa} \right)
\end{equation}

\noindent
where $w_k$ is a weight associated with patient speciality type $k$. For this
study we assume $w_k = 1$ for all $k$, which allows the interpretation of
$g\left(Z_a, \tilde{Z}_a\right)$ as the expected number of patients surviving
per time unit. Now $\Psi_{kpa}$ can be interpreted as
the probability of a patient $k \in \mathcal{K}_B$ at pick-up location $p$
being seen by a primary vehicle from location $a$ and surviving; while
$\hat{\Psi}_{kpa}$ would be the probability of patients of speciality
$k \in \mathcal{K}_A$ at pick-up location $p$ being seen by any vehicle from
location $a$ and surviving. Equation~\ref{eqn:meslmhphf} is the weighted sum
over the expected survival probabilities of all patient specialities, all
patient pick-up locations, and all ambulance stations. These survival
probabilities are given by Equations~\ref{eqn:survival_A}
and~\ref{eqn:survival_B} respectively.

\begin{equation}\label{eqn:survival_A}
\Psi_{kpa} = s_k\left( b_{pa} \right)
\left(1 - \pi_{a}^{Z_a} \right)
\prod_{\alpha \in \mathcal{A}}
\pi_{\alpha}^{\left(Z_{\alpha} \beta_{p\alpha a} \right)}
\end{equation}

\begin{align}\label{eqn:survival_B}
\hat{\Psi}_{kpa} &= s_k\left(\tilde{b}_{pa}\right)
\left(1 - \tilde{\pi}_{a}^{\tilde{Z}_a} \right)
\prod_{\alpha \in \mathcal{A}}
\tilde{\pi}_{\alpha}^{\left(\tilde{Z}_{\alpha} \beta_{p\alpha a}\right)}
\pi_{\alpha}^{\left(Z_{\alpha} R_{p \alpha a}\right) } \nonumber \\
&+ s_k\left(b_{pa}\right) \left(1 - \pi_{a}^{Z_a} \right)
\prod_{\alpha \in \mathcal{A}}
\pi_{\alpha}^{\left(Z_{\alpha}\beta_{p\alpha a}\right)}
\tilde{\pi}_{\alpha}^{\left(\tilde{Z}_{\alpha}
\left(1 - R _{p a\alpha}\right)\right)}
\end{align}

Here $R_{p a_1 a_2}$ and $\beta_{p a_1 a_2}$ \textcolor{red}{define an ordering
of the ambulance locations, they} are binary variables indicating the preference
of sending a vehicle from one station to another, that is the dispatch rule.
Here we use the closest or fastest vehicle, so $\beta_{p a_1 a_2}$ indicates if
a vehicle of the same type can reach $p$ quicker from $a_1$ than $a_2$, defined
in Equation~\ref{eqn:beta}, while $R_{p a_1 a_2}$ indicates if a primary vehicle
at $a_1$ can reach $p$ quicker than a secondary vehicle at $a_2$, defined in
Equation~\ref{eqn:R}. \textcolor{red}{In order for these to define an ordering
it is assumed that no two distinct ambulance locations are exactly equidistant
from a pickup location.} Note that here these parameters are defined using
travel times, but can be generalised to account for any type of preference,
including for example proximity, importance, or efficiency.

\begin{equation}\label{eqn:beta}
\beta_{p a_1 a_2} = \begin{cases}
    0 & \text{if } a_1 = a_2\\
    1 & \text{if } b_{p a_1} < b_{p a_2}\\
    0 & \text{otherwise.}
\end{cases}
\end{equation}

\begin{equation}\label{eqn:R}
R_{p a_1 a_2} = \begin{cases}
    1 & \text{if } b_{p a_1} \leq \tilde{b}_{p a_2}\\
    0 & \text{otherwise.}
\end{cases}
\end{equation}

Equation~\ref{eqn:survival_A} is the probability of a patient surviving (their
survival function), multiplied by the availability of primary vehicles at that
ambulance station, multiplied by the unavailability of vehicles at closer
stations.
Equation~\ref{eqn:survival_B} extends this to two vehicles types, the first
part of the sum repeating the logic for the faster secondary vehicles, and the
second part of the sum adapting that logic for primary vehicles, who will only
be contribute the patient's survival if all faster secondary vehicles are also
busy. These interpretations are given as annotations to the equations
in~\ref{apx:annotated}.

A key consideration is the vehicle utilisations $\pi_a$ and $\tilde{\pi}_a$;
discussed in the next subsection.

\subsection{Utilisation Considerations}\label{sec:utilisation}
The models above assume that the utilisations of each vehicle type at each
vehicle location is known, which is a potentially restrictive assumption.
These utilisations $\pi_a$ and $\tilde{\pi}_a$ depend on the demand to station
$a$, which itself depends on the vehicle allocations. One method of overcoming
this, used in \citet{Knight2012918}, is to consider utilisation as the ratio of
demand and service rates, shown in Equations~\ref{eqn:utilisation_ratio_primary}
and~\ref{eqn:utilisation_ratio_secondary}. Here we let $\frac{1}{\mu}$ and
$\frac{1}{\tilde{\mu}}$ be the average job times of primary and secondary
vehicles, then $\lambda_a$ and $\tilde{\lambda}_a$ represent the share of the
demand seen by primary and secondary vehicles from vehicle location $a$,
respectively.

\begin{align}
\pi_a &= \frac{\lambda_a}{\mu} \label{eqn:utilisation_ratio_primary}\\
\tilde{\pi}_a &= \frac{\tilde{\lambda_a}}{\tilde{\mu}} \label{eqn:utilisation_ratio_secondary}
\end{align}

Note that here we assume that average job times are not dependent on the vehicle
location $a$, although this may be unrealistic, given that the travel between
locations are a key part of a vehicle's job time ($B_{pa}$, $D_{ya}$, $F_{pa}$).
However, as discussed in Section~\ref{sec:problem_description}, there are a
number of other components to an ambulance job including time on site ($G_k$),
hospital handover time ($J_k$), and vehicle refill time ($\Theta$). For the
case of Jakarta, we justify the assumption of job times not being location
dependent by considering the proportion of an ambulance job's time that is
location dependent, $\rho$, defined in
Equation~\ref{eqn:proportion_location_dependant},

\begin{equation}\label{eqn:proportion_location_dependant}
\rho = \begin{cases}
  \frac{B_{pa} + D_{ya}}{T_{paky}} & \text{ if hospital required,} \\[8pt]
  \frac{B_{pa} + F_{pa}}{T_{pak}} & \text{ if no hospital required,}
\end{cases}
\end{equation}

\noindent
then, using the simulation described in Section~\ref{sec:simulation}, we
consider the distribution of $\rho$ across all simulated jobs under the current
situation and allocation in Jakarta, shown in Figure~\ref{fig:rho_distribution}.
We see for primary vehicle jobs on average 24\% of a job time is location
dependent, while it is 36.5\% for secondary vehicles. Considering
location-dependent service rates would be an interesting further research
question, potentially improving the model's performance by incorporating travel
times into the MESLMHPHF. However, given that in the case of Jakarta the
preferences $\beta_{p a_1 a_2}$ and $R_{p a_1 a_2}$ are themselves defined by
travel times (Equations~\ref{eqn:beta} and~\ref{eqn:R}), this may not have much
affect for this study, \textcolor{red}{as the model already gives preference to
the locations with the shortest service times.}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{img/location_dependant_service_time_proportion}
    \caption{Distribution of the proportion of an ambulance job consists of
    location dependent components, for primary and secondary vehicles.}
    \label{fig:rho_distribution}
\end{figure}

The demand experienced by vehicles at each location $a \in \mathcal{A}$ would
indeed be location specific, and would depend on the number and utilisation of
the vehicles at every other location. This is addressed in the next subsection.

\subsection{Numerical approximation of utilisation}
For primary vehicles, which operate independently of secondary vehicles, the
relationship between demand experienced at each location, $\lambda_a$, and the
allocations and utilisations of all other locations, is given by
Equation~\ref{eqn:utilisations_primary}. For secondary vehicles, the
relationship between demand experienced at each location, $\tilde{\lambda}_a$,
and the allocations and utilisations of all other locations, is given by
Equation~\ref{eqn:utilisations_secondary}. Note that this also depends on the
utilisations of the primary vehicles, $\pi_a$.

\begin{equation}\label{eqn:utilisations_primary}
\lambda_a = \sum_{p \in \mathcal{P}} \sum_{k \in \mathcal{K}} \lambda_{pk}
\left( 1 - \left(\frac{\lambda_a}{\mu}\right)^{Z_a} \right)
\prod_{\alpha \in \mathcal{A}} \left(\frac{\lambda_{\alpha}}{\mu}\right)^{
\left(Z_{\alpha} \beta_{p \alpha a}\right)}
\end{equation}

\begin{equation}\label{eqn:utilisations_secondary}
\tilde{\lambda}_a = \sum_{p \in \mathcal{P}}
\sum_{k \in \mathcal{K}_A} \lambda_{pk}
\left( 1 - \left(\frac{\tilde{\lambda}_a}{\tilde{\mu}}\right)^{\tilde{Z}_a} \right)
\prod_{\alpha \in \mathcal{A}} \pi_{\alpha}^{\left(Z_{\alpha} R_{p \alpha a}\right)}
\left(\frac{\tilde{\lambda}_{\alpha}}{\tilde{\mu}}\right)^{
\left(\tilde{Z}_{\alpha} \beta_{p \alpha a}\right)}
\end{equation}

We propose finding the true vehicle utilisations by first solving
Equation~\ref{eqn:utilisations_primary} for the $\lambda_a$; determining the
primary vehicle utilisations with Equation~\ref{eqn:utilisation_ratio_primary};
then using these to solve Equation~\ref{eqn:utilisations_secondary} for the
$\tilde{\lambda}_a$; and determining the secondary vehicle utilisations with
Equation~\ref{eqn:utilisation_ratio_secondary}. These can be solved
numerically. In our implementation this is solved using the MINPACK hybrd and
hybrj algorithms by using Scipy's \texttt{fsolve} function
\citep{2020SciPy-NMeth}. It should be noted that in this section we assume that
all $\lambda_{pk}$ are static and not time-dependant, which is an assumption
that will be used throughout the optimisation methodology, however in the
simulation we can account for time-dependent demand.

Now note that, in Equation~\ref{eqn:utilisations_primary}, we have that

$$\sum_{p \in \mathcal{P}} \sum_{k \in \mathcal{K}} \left( 1 - \left(\frac{\lambda_a}{\mu}\right)^{Z_a} \right) \prod_{\alpha \in \mathcal{A}} \left(\frac{\lambda_{\alpha}}{\mu}\right)^{\left(Z_{\alpha} \beta_{p \alpha a}\right)} < 1,$$

\noindent
and so
$\sum_{a \in \mathcal{A}} \lambda_a < \sum_{p \in \mathcal{P}} \sum_{k \in \mathcal{K}} \lambda_{pk}$,
meaning that there is lost demand. This corresponds to the patients who are
abandoned, or take private or public transport to a hospital instead of waiting
for an ambulance. Similarly, in Equation~\ref{eqn:utilisations_secondary}, we
have that

$$\sum_{p \in \mathcal{P}} \sum_{k \in \mathcal{K}_A} \left( 1 - \left(\frac{\tilde{\lambda}_a}{\tilde{\mu}}\right)^{\tilde{Z}_a} \right) \prod_{\alpha \in \mathcal{A}} \pi_{\alpha}^{\left(Z_{\alpha} R_{p \alpha a}\right)} \left(\frac{\tilde{\lambda}_{\alpha}}{\tilde{\mu}}\right)^{ \left(\tilde{Z}_{\alpha} \beta_{p \alpha a}\right)} < 1,$$

\noindent
and so
$\sum_{a \in \mathcal{A}} \tilde{\lambda}_a < \sum_{p \in \mathcal{P}} \sum_{k \in \mathcal{K}_A} \lambda_{pk}$.
This lost demand represents both abandoned calls, but also calls where secondary
vehicles are not deployed as a primary vehicle could reach them first.

\subsection{Metaheuristic Optimisation}\label{sec:heuristic}
In order to maximise the MESLMHPHF objective function presented in
Section~\ref{sec:meslmhphf}, we use an evolutionary algorithm. A population of
possible solutions is created and ranked according to the objective, then for
each generation of the metaheuristic a proportion of the best performing
solutions is kept, and are mutated to complete the population for the next
generation. To encourage exploration, the number of times each solution is
mutated begins high and gradually decreases throughout the run of the
simulation. The metaheuristic takes five hyper-parameters: $N$ the population
size, $\kappa$ the number of solutions to retain to the next generation, $m_0$
the initial mutation rate; $c$ the cooling rate; and $H$ the number of
generations. The metaheuristic is described in Algorithm~\ref{alg:heuristic}.

\begin{algorithm}
\algnewcommand\algorithmicto{\textbf{to }}
\caption{Evolutionary Algorithm used to find better allocations.}
\label{alg:heuristic}
\begin{algorithmic}[1]
\State Create a population of $N$ random allocations ($Z_a$, $\tilde{Z}_a$)\;
\For{$i \leftarrow 0$ \algorithmicto $H$}
  \State Rank the population \textcolor{red}{by decreasing expected survival} according to Equation~\ref{eqn:meslmhphf}\;
  \State Keep the top $\kappa$ solutions\;
  \State $m \leftarrow \lceil m_0 - \max\left(m_0, i \times c\right)\rceil$\;
  \For{$j \leftarrow 0$ \algorithmicto $N - \kappa$}
    \State Randomly choose and copy a solution from the population\;
    \State Mutate that copy of the solution $m$ times\;
    \State Place mutated copy in a separate population\;
  \EndFor
  \State Combine populations;\
\EndFor
\State Rank the population according to Equation~\ref{eqn:meslmhphf};\
\State Output the top ranking solution.
\end{algorithmic}
\end{algorithm}

Mutations consist of one of four possible changes: randomly change the location
of one primary vehicle; randomly change the location of one secondary vehicle;
randomly remove one primary vehicle and add secondary vehicles to three randomly
chosen locations (as one primary vehicle costs the equivalent to three secondary
vehicles); and randomly remove three secondary vehicles and add a primary
vehicle to a randomly chosen location. If it is required that exact vehicle
numbers need to be retained, then only for first two mutations need be chosen.

The source code for the optimisation metaheuristic, including the evaluation of
the MESLMHPHF objective function and the solution method for the utilisation
considerations, is available at \citet{zenodo_opt} and development took place at
\url{https://github.com/drvinceknight/HeterogeneousAmbulanceFleetAllocations}.
Numerical results are discussed in Section~\ref{sec:jakarta}.

\section{Simulating Allocations}\label{sec:simulation}
Discrete event simulation (DES) is a common methodology that allows us to
investigate given scenarios under uncertainty. It is a common methodology for
modelling emergency medical services, with a review given in
\citet{aboueljinane13}. Here it will be used to quantify the effectiveness of
given ambulance allocations by measuring a range of key performance indicators
(KPIs), such as the average response time, the percentage of abandoned calls,
vehicle utilisations, and expected survival based on response times. The
simulation is built using the Ciw library \citep{palmer2019ciw}, and the model
logic is described below. Its central ideas include modelling transit jobs as
customers, rather than the patients themselves, and simulating primary and
secondary vehicles as two simulations sequentially, with the output of the
former being the input of the latter. Separating out primary and secondary
vehicle logic simplifies the model logic, allows for modular simulations that
can be run in isolation, and so easier to adapt and to maintain.

\subsection{Simulating Primary Vehicles}\label{sec:simulation_primary}
For primary vehicles the logic to simulate is as follows: Patients make a call
from one of a number of pick-up locations and await an ambulance to pick them
up and take them to an appropriate hospital. Ambulances are stationed at a
number of ambulance locations; when a patient makes a call, all free
ambulances from any location calculate their expected time to reach that
patient, and the ambulance with the smallest expected time to the patient is
called out. The ambulance drives from its current ambulance location to the
patient's pick-up location, then if a hospital is required, from the patient's
pick-up location to the hospital, and then from the nearest hospital back to
their original ambulance location. If a hospital is not required, the
ambulance returns to their original ambulance location.

Rather than considering patients as customers in a queue, here the situation
is re-framed to model transit jobs as customers, with the ambulances as
servers. This is similar to the approach of \citet{kergosien15}, whereas
ambulances are now simple servers, their stations are nodes in a queueing
network, with a routing decision representing the ambulance preference.

Transit jobs can be categorised into classes corresponding to the pick-up
locations $\mathcal{P}$ and speciality $\mathcal{K}$. Jobs of class
$(p, k) \in \mathcal{P} \times \mathcal{K}$ arrive with rate $\lambda_{pk}$.
Servers can be categorised into classes corresponding to the ambulance
locations $\mathcal{A}$. Service times are server-dependent, that is the
service time of a transit job of type $(p, k)$, being served by a server of
class $a$, will have service time $T_{paky}$ if a hospital is required and
$T_{pak}$ if no hospital is required. These were given by
Equations~\ref{eqn:service_time_withhosp} and~\ref{eqn:service_time_nohosp}
respectively, in Section~\ref{sec:problem_description}.
That is $T_{paky}$ and $T_{pak}$ is the overall time the ambulance spends
dealing with that transit job and is unavailable to receive any more transit
jobs.

From a patient's point of view their service only lasts $G_k + C_{py} + J_k$,
(or $G_k$ if no hospital is required), and their waiting time is $B_{pa}$ plus
the time waiting for an ambulance to be dispatched. However in our case, if
there are no free ambulances at the time of call, it is assumed that patients
find their own care or transport, and so the call is abandoned. Therefore, the
time waiting for dispatch is always zero.

Furthermore, all service times are time dependent, and calculated from travel
distances and approximate hourly traffic levels, described in
Section~\ref{sec:travel_times}.

\subsection{Travel Time Calculations}\label{sec:travel_times}
Travel times within a city such as Jakarta are not constant throughout the day
due to the variability in traffic, and it is necessary to capture these in
models of emergency medical services \citep{schmiddoerner10}.
Here, each day is split into a set of periods, $\mathcal{H}$ indexed by $h$.
Within each period traffic levels are modelled as piecewise linear functions
similar to \citet{horn00}, and are here considered constant but differing from
period to period. Traffic levels influence the speed of the ambulance through a
delay factor $d_h$ associated with each period. Therefore, the expected speed
$S_h$ at which an ambulance travels during period $h$ is given by
Equation~\ref{eqn:speed},

\begin{equation}\label{eqn:speed}
S_h = s d_h
\end{equation}

\noindent
where $s$ is some given baseline speed. The relationship between time travelled
and distance covered is piece-wise linear with slope $S_h$ in period $h$.
Therefore, travel times are calculated using this relationship.

As an example, consider the scenario where for the first period $t \in (0, 4)$
we have $S_1$ = 1; for $t \in (4, 8)$ we have $S_2 = \frac{1}{4}$, for
$t \in (8, 12)$ we have $S_3 = 2$, and for $t \in (12, 16)$ we have $S_4 = 1$.
Consider that the vehicle must travel 11 units. If the vehicle begins its
journey at $t=0$, then the expected travel time would be 11 time units, shown
in Figure~\ref{fig:travel_times_1}. However if the vehicle begins its journey
at time $t=3$ then the expected travel time is 10 time units, shown in
Figure~\ref{fig:travel_times_2}.

\begin{figure}
    \begin{center}
    \begin{subfigure}{6.6cm}
    \includestandalone[scale=0.33]{img/travel_times}
    \caption{Beginning at $t=0$, a travel time of $11$.}
    \label{fig:travel_times_1}
    \end{subfigure}
    \begin{subfigure}{6.6cm}
    \includestandalone[scale=0.33]{img/travel_times_2}
    \caption{Beginning at $t=3$, a travel time of $10$.}
    \label{fig:travel_times_2}
    \end{subfigure}
    \end{center}
    \caption{Example of calculating travel times for two different starting
    times.}
    \label{fig:travel_times}
\end{figure}

This is the method used to calculate the expected values of $B_{pa}$, $C_{py}$,
$D_{ya}$ and $F_{pa}$.
Furthermore it is assumed that travel times for each segment of the route
(from vehicle location to patient, from patient to hospital, and from hospital
back the the vehicle's location) follow a Triangular distribution around the
calculated expected travel time, between 75\% and 125\% of the value.

\subsection{Time-dependent Demand}
Additionally, it is assumed that calls arrive according to a Poisson
distribution, with rates $\lambda_{pk}$, and that these rates are time
dependent. Each day is split into four 6-hour periods, the morning, afternoon,
evening and night, and each speciality $k$ and pick-up location $p$ will have
a different call arrival rate for each of these periods.

For some locations and specialities the number of observed calls might be very
low, and so the $\lambda_{pk}$ rate would be very low. This can cause
synchronicity issues when sampling arrivals \citep{pidd2004computer}, where
artificially long inter-arrival times can be introduced at the beginning of
one time period due to the low arrival rate in the previous time period.
In order to overcome this here, rather than sample inter-arrival times
iteratively from an Exponential distribution, an entire schedule of arrival
dates are sampled at the beginning of the simulation run, first by sampling
the number of arrivals required in each time period from a Poisson
distribution, and then by sampling specific dates within that time period
using a Uniform distribution.
This mechanism, and alternative to thinning \citep{lewisshedler79}, was
implemented in the Ciw software in version v2.3.3 \citep{ciw233}, and described
in the documentation here:
\url{https://ciw.readthedocs.io/en/latest/Guides/time_dependent.html}.

\subsection{Simulating Secondary Vehicles}\label{sec:simulation_secondary}
Secondary non-transit vehicles, can in general travel faster than primary
transport vehicles, and so can be dispatched at the same time as a primary
vehicle but reach the patient earlier, reducing the response time for that
patient and so increasing their survival probability. They are only dispatched
for patients of speciality $k \in \mathcal{K}_A$, and if the closest secondary
vehicle can reach them before the closest primary vehicle.
Here secondary vehicles are simulated in a second discrete event simulation,
run sequentially, but simulating the exact same time period and events as the
first simulation. This is similar to sequential hybrid simulation methodology
\citep{brailsfordetal19, morganetal17}, however in this case the two combined
components are both DES, and are combined to simplify the logic of each
component, maintain modularity and so ease model reusability and future
adaptations.

This is possible as there are no synchronicity issues between the two
components. Primary vehicles operate independently of secondary vehicles, that
is the way primary vehicles respond to a patient is unaffected by the presence
or lack of a secondary vehicle.
Therefore, primary vehicle logic is not compromised by simulating primary
vehicles in isolation.
Secondary vehicles are impacted by the behaviour of primary vehicles, they
must remain with the patient until the primary vehicle arrives, and so must be
simulated after the simulation of primary vehicles, taking as inputs the exact
list of events that occurred. That is the logic of the secondary vehicles is
determined by observing the actions of primary vehicles and reacting to them.
Simulation results are combined for each individual patient, to determine
their response time caused by either primary or secondary vehicles.

Secondary vehicles are chosen in the same way as primary vehicles, by choosing
out of the currently free vehicles the one with the smallest expected time to
patient. Their service times are reactionary to what occurred with the primary
vehicle. The decision to send a primary vehicle or not depends on the previously
estimated travel time for the primary vehicle, not the actual travel time
experienced by that primary vehicle: if a secondary vehicle is estimated to
reach there first, they will be dispatched.
Let $\tilde{B}_{pa_2}$ be the time it takes for the secondary vehicle to travel
from its location $a_2$ to the patient pick-up location $p$; $B_{pa_1}$ is the
time is took for the primary vehicle to get from its location $a_1$ to the
patient pick-up location $p$; $G_k$ is the time the primary vehicle spent
with the patient; and $\tilde{F}_{pa_2}$ is the time it takes to return to the
secondary vehicle's location $a_2$ from the patient pick-up location $p$.
Note that $B_{pa_1}$ and $G_k$ are exact values obtained from the
initial simulation of the primary vehicles, while $\tilde{B}_{pa_2}$ and
$\tilde{F}_{pa_2}$ are random variables to sample in the subsequent simulation.
Travel times are calculated as described in Section~\ref{sec:travel_times},
replacing the primary vehicle delay factor, $d_h$, with a delay factor for
secondary vehicles, $\tilde{d}_h$. This accounts for secondary vehicles
travelling faster and reacting to traffic differently to primary vehicles,
similar to the method used in \citet{hendersonmason04} to differentiate the
travel times of ambulances with and without flashing lights.

Exact synchronicity considerations are shown in Figure~\ref{fig:sequential_logic}:
\begin{itemize}
  \item whenever the secondary vehicle reaches the patient before the primary
  has left, vehicle, they remain with the patient until the primary vehicle
  leaves;
  \item if secondary vehicle arrives after the primary vehicle has left, they
  will immediately return to their stations as they are not needed at the scene;
  \item if the expected time for the secondary vehicle to reach the patient
  exceeds the expected time for the primary vehicle to reach the patient then
  the secondary vehicle is not deployed, and so that transit job would be
  abandoned in the secondary simulation (although still seen by a primary
  vehicle in the initial simulation). We assume that secondary vehicles must
  still reach the pickup locations, due to potential difficulties in
  communication en-route.
\end{itemize}

Therefore job service times for secondary vehicles, $\tilde{T}_{pak}$, are
given by:

\begin{figure}
    \centering
    \includestandalone[scale=0.6]{img/synchronicity}
    \caption{Visualising secondary vehicle (SV) logic reacting to the primary
    vehicle's (PV) actions.}
    \label{fig:sequential_logic}
\end{figure}

\begin{equation}
  \tilde{T}_{pak} = \max\left(\tilde{B}_{pa_2}, B_{pa_1} + G_k \right) + \tilde{F}_{pa_2}
\end{equation}

\subsection{Combined Model}
Combining the logic presented in Sections~\ref{sec:simulation_primary}
and~\ref{sec:simulation_secondary} in a sequential manner, with the output of
one determining the input of the other, gives the overall simulation logic for
simulating both primary and secondary vehicle types. It is used to quantify
the effectiveness of a given allocation $(Z_a, \tilde{Z}_a)$ for all
$a \in \mathcal{A}$, by recording several useful KPIs.
In this work, the KPIs of interest are: the average primary vehicle
utilisation, the average secondary vehicle utilisation, the mean response
time, the percentage of abandoned calls (that is a measure of the primary
vehicle unavailability), and the expected overall survival. Survival is
modelled using a combination of survival function curves and hard cut-offs,
and is described in Section~\ref{sec:survival}.
The source code for the combined simulation, is available at \citet{zenodo_sim} and
development took place at \url{https://github.com/MarkTuson/ambulance_simulation}.

\section{The Case of Jakarta}\label{sec:jakarta}
Jakarta, the capital of Indonesia, has a population of 10.5 million
\citep{BPS_Jakarta} residing within 664 km$^2$. The city is divided into five
municipalities and one district, each of which is divided into sub districts
and, in turn, neighbourhoods. There are in total 42 sub districts and 261
neighbourhoods within mainland Jakarta (excluding the Thousand Islands regency
in the north) \citep{BPS_Jakarta_angka}; a map is given in
Figure~\ref{fig:region_names}.

As described in Section~\ref{sec:intro}, there are many challenges to calling
for an ambulance in Jakarta, as captured in \citet{BriceSyaribahNoor2022Esui}.
Based on this work the regional government of Jakarta invested in a new fleet
of coordinated ambulances accessible via the single emergency number 119.
This work, in collaboration with Ambulans 118 and the 119 Emergency Ambulance
Service, finds potential fleet allocations using the optimisations methods
given in Section~\ref{sec:optimising}, and evaluates these allocations'
effectiveness using the simulation described in Section~\ref{sec:simulation}.

\begin{figure}
\begin{center}
\includegraphics[width=0.4\textwidth,trim={0 6cm 0 5cm}, clip]{img/jakarta_region_names.pdf}
\end{center}
\caption{Map of Jakarta's Municipalities.}
\label{fig:region_names}
\end{figure}

As of October 2022, the 119 service in Jakarta ran a fleet of 81 primary
Emergency Ambulances (EAs) and 13 secondary motorbike Rapid Response Vehicles
(RRVs), distributed across 67 ambulance stations throughout the city.
Figure~\ref{fig:current_allocation} summarise the allocations vehicle
locations.
Working with our ambulance partners, after careful consideration the following
patient `specialities' ($k \in \mathcal{K}$) were agreed based on clinical need
(although other categories could readily be adopted and included as necessary):

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{img/map_current}
\caption{Map of current Emergency Ambulance (EA) and Rapid Response Vehicle
         (RRV) locations and allocations to ambulance bases across Jakarta.}
\label{fig:current_allocation}
\end{center}
\end{figure}

\begin{itemize}
  \item \textbf{A1 - High priority emergency patients}: critical patients who
  require immediate life-saving assistance with a target ambulance response
  time of 8 minutes. In our data set there were 168 identified calls that met
  this criterion.
  \item \textbf{A2 - Other emergency patients}: urgent patients that require
  assistance with a target response time of 15 minutes. In our data set there
  were 23,784 calls of this type.
  \item \textbf{B - Non-emergency patients}: patients that still benefit from
  an ambulance response but non-critical with a target response time of 60
  minutes. There were 31,725 calls of this type in the data set.
\end{itemize}

Secondary vehicles can be called to patients of speciality A1 and A2,
therefore $\mathcal{K}_A = \{\text{A1}, \text{A2}\}$ and
$\mathcal{K}_B = \{\text{B}\}$.
The number of calls from each of these specialities varies by neighbourhood,
as shown in Figure~\ref{fig:yearly_demand}. It can be seen that many calls are
highly concentrated in a handful of neighbourhoods rather than spread
throughout the city. In this work, we approximate pick-up locations by the
geographic centroid of each of the 261 neighbourhoods, each representing an
ambulance pick-up that that occurred within that neighbourhood.

\begin{figure}
\begin{center}
\includegraphics[width=0.85\textwidth]{img/yearly_demand.pdf}
\end{center}
\caption{Number of calls received by speciality and neighbourhood.}
\label{fig:yearly_demand}
\end{figure}

Firstly, in Section~\ref{sec:analysis_grid} we use the MESLMHPHF objective
function of expected survival, and the simulation, to find KPIs that to compare
the performance of two proposed grid allocations to the currently used
allocation. An additional KPI is also calculated, the expected survival of A1
patients, found by taking Equation~\ref{eqn:meslmhphf} and summing over
$k = \text{A1}$ only.
Then in Section~\ref{sec:demand_scenarios} we outline four possible
future demand scenarios, and in Section~\ref{sec:optimise_current} we use the
optimisation model to find and compare improved allocations for the current
number of vehicles. Finally in Section~\ref{sec:vehicle_numbers} we investigate
the number of vehicles required to meet the demand across the given scenarios.

All models are parameterised using demand data that covers all 261
neighbourhoods from 1 January to 31 December 2019, before the COVID-19
pandemic. Data from 2020-2021 was naturally heavily skewed by the pandemic,
with significantly lower demand, and so was not considered as representative
of a typical period for forecasting future needs, hence the decision to use
the available 2019 demand. \ref{apx:parameters} provides further details on the
parameterisation of the model, including how travel time estimates were
obtained.

\subsection{Evaluating `Grid' Proposals}\label{sec:analysis_grid}
Discussions with senior staff at 119 identified that they were seriously
considering a re-configuration their ambulance allocations into a grid
structure, placing an ambulance station at regular intervals throughout the city
and uniformly distributing the vehicles that they felt would ensure equitable
coverage. Two possible grid allocations were being considered: one placing an
EA every 3km across the city (giving a total of 70 vehicles), and another
placing three EAs every 5km across the city (giving a total of 72 vehicles).
Figure~\ref{fig:grid_allocations} show these proposed allocations.

\begin{figure}
\begin{center}
\begin{subfigure}{0.37\textwidth}
\includegraphics[width=\textwidth]{img/map_grid3km_proposed}
\caption{3km grid.}
\end{subfigure}
\hspace{8mm}
\begin{subfigure}{0.37\textwidth}
\includegraphics[width=\textwidth]{img/map_grid5km_proposed}
\caption{5km grid.}
\end{subfigure}
\caption{Proposed grid allocations.}
\label{fig:grid_allocations}
\end{center}
\end{figure}

Running the simulation and MESLMHPHF objective function of expected survival for
both the current and the two proposed grid allocations gives the results shown
in Table~\ref{tbl:current_grid_results}. The simulation was run over half a
year, with a month warm up time, over eight replications. We observe that the
two proposed grid allocations perform much worse than the current allocation as
measured by mean response time, survival, and percentage of calls abandoned.
This is expected as vehicle numbers are lower under the new proposals, as well
as the allocations focussing on coverage rather than response times or survival.

\begin{table}
\begin{tabular}{rrrr}
\toprule
Allocation & Baseline & Grid 3km & Grid 5km \\
\midrule
Number of EAs & 81 & 70 & 71 \\
Number of RRVs & 13 & 0 & 0 \\
Ambulance Utilisation & 30.99\% & 36.58\% & 35.79\% \\
RRV Utilisation & 22.47\% & - & - \\
Mean Response Time (mins) & 17.69 & 22.29 & 24.05 \\
MESLMHPHF Objective & 98.34\% & 87.95\% & 85.64\% \\
Expected A1 Survival & 26.05\% & 13.64\% & 12.23\% \\
Percent Abandoned & 0.00\% & 0.52\% & 0.93\% \\
\bottomrule
\end{tabular}
\caption{Calculated KPIs for the current and proposed grid allocations.}
\label{tbl:current_grid_results}
\end{table}

\subsection{Planning for Future Demand}\label{sec:demand_scenarios}
Similarly to many other cities worldwide, ambulance services in Jakarta
anticipate that demand for services will grow in future years, not least because
there is now a coordinated service accessible via a single common number `119'
to call. This should help raise awareness and visibility of an ambulance
service. In the case of Jakarta, certainly increased use of 119 and therefore
increased demand would be a desired outcome and is pro-actively supported
by the Indonesian Government.

We consider four different demand scenarios, developed by considering the
result of our cross-sectional study of patients attending EDs in Jakarta
\citep{BriceSyaribahNoor2022Esui}. As part of the survey, randomly selected
patients arriving at each emergency department were asked whether they had
used an ambulance to attend, and reasons for not using an ambulance. For those
patients who were categorised by the medical staff as emergency and for whom
therefore an ambulance might have seemed a sensible option, the responses are
given in Table~\ref{table:survey_results}.

\begin{table}
\centering
\begin{tabular}{rr}
\toprule
Reason & \% of emergency respondents \\
\midrule
Used Ambulance & 13.0\\
Too expensive & 5.2  \\
Not available  & 13.1 \\
Would take too long & 16.5 \\
Not necessary & 15.7  \\
Not aware of service & 33.1\\
Other &3.5 \\
\bottomrule
\end{tabular}
\caption{Barriers to use of an ambulance service in Jakarta, from
         \citet{BriceSyaribahNoor2022Esui}.}
\label{table:survey_results}
\end{table}

The survey confirmed three major barriers to use, namely: the service's
visibility (33.1\% of respondents were unaware of the service), the service's
reliability (13.1\% of respondents reported that there was no ambulance
available and 16.5\% of respondents believed it would take too long), and the
service's cost (5.2\% of respondents cited cost as a deterrent). Therefore,
four demand scenarios were considered, corresponding to addressing each of
these barriers in turn:

\begin{itemize}
  \item \textbf{D13}: representing the current situation where approximately
        13\% of emergency patients (specialities A1 and A2) do use an ambulance.
  \item \textbf{D19}: representing the situation where visibility is addressed.
        Here, we distribute the 33.1\% of the respondents who were unaware of
        the service proportionally between using the ambulance and amongst the
        remaining issues. Using this methodology we would expect 19.4\% of
        emergency patients to now use an ambulance.  
  \item \textbf{D34}: representing the situation where reliability and
        visibility are addressed. Using the same methodology we would expect
        34.8\% of emergency patients to now use an ambulance.
  \item \textbf{D45}: representing the situation where cost, reliability and
        visibility are all addressed. Using the same methodology we would expect
        45.4\% of emergency patients to now use an ambulance. 
\end{itemize}

Guided by our findings in \citet{BriceSyaribahNoor2022Esui}, recent changes
mean that the 119 services is now free to use for Jakarta residents, so
certainly the highest demand (\textbf{D45}) scenario is plausible in the near
future. After discussions with our partners, it was agreed that for each of
the scenarios we should assume that non-emergency demand (speciality B)
remains unchanged.

\subsection{Improving the Current Allocation}\label{sec:optimise_current}
Applying the optimisation metaheuristic from Section~\ref{sec:optimising} to the
current allocation re-allocates the 81 EAs and 13 RRVs. We do this for each of
the four scenarios described in Section~\ref{sec:demand_scenarios}. The
hyper-parameters used for this optimisation algorithm throughout this study are
given in Table~\ref{tbl:hyperparameters}, which include approximations for the
average service times of primary and secondary vehicles derived from the initial
simulation model. \ref{apx:hyperparameters} shows some initial explorations on
the hyper-parameter choices, giving confidence that the chosen parameters, with
a large enough number of iterations $N$, are sufficient to find allocations
that perform well.

\begin{table}
\centering
\begin{tabular}{lll}
\toprule
Parameter & Value Used & Explanation \\
\midrule
$\mu$ & $\frac{1}{3.886 \times 60}$ & Average primary vehicle service rate ($\text{hr}^{-1}$).\\
$\tilde{\mu}$ & $\frac{1}{1.038 \times 60}$ & Average secondary vehicle service rate ($\text{hr}^{-1}$).\\
$N$ & 100 & Population size.\\
$\kappa$ & 20 & Number of solutions to keep per generation.\\
$m_0$ & 6 & Initial number of mutations.\\
$c$ & 0.1 & Cooling rate, rate at which number of mutations decreases.\\
$H$ & 200 & Number of generations.\\
\bottomrule
\end{tabular}
\caption{Hyper-parameters used in the optimisation for all experiments.}
\label{tbl:hyperparameters}
\end{table}

Table~\ref{tbl:current_optimal_compare} compares the MESLMHPHF objective
function value and simulation derived KPIs between the current and the optimised
allocations, for each demand scenario. It can be seen that, as expected, as
demand increases then utilisation of primary and secondary vehicles increase, as
the ambulance service is busier. Similarly, mean response time increases with
demand, this is due to the vehicles being busier, and therefore less chance that
the most optimally placed vehicle is dispatched. This is confirmed by the
percentage of calls abandoned increasing with demand, showing ambulance
unavailability.
Looking at the expected survival, that is the value of the MESLMHPHF objective
function, we see that using an better allocation does increase the survival,
and that the drop in survival due to demand increases is not as severe when
using an improved allocation in comparison to the current allocation. We can
also see that in particular A1 patients are benefiting from the improved
allocations, as these are the patients most sensitive to small fluctuations in
response time, due to their survival function Equation~\ref{eqn:survival_curve}.
It is worth noting that for the simulation based KPIs (mean response times and
percent abandoned), there is not too much
difference between the currently used allocation and the better allocation,
showing that these KPIs may not be good approximations of survival.

\begin{table}
\resizebox{\textwidth}{!}{%
\input{tbl/current_vs_optimised.tex}%
}
\caption{Calculated KPIs for the current and improved allocations under the
four possible demand scenarios.}
\label{tbl:current_optimal_compare}
\end{table}

\subsection{Finding Better Allocations for Any Resource Level}\label{sec:vehicle_numbers}
The metaheuristic algorithm described in Section~\ref{sec:heuristic} finds
allocation of ambulances across the 67 current ambulance stations for a given
number of primary and secondary vehicles. Information supplied by one of the
ambulance operators suggested that in terms of total running costs one EA
(primary vehicle) was approximately equivalent to three RRVs (secondary
vehicles), and so we consider three RRVs to be one resource; that is, a
resource level of 60 could represent 60 primary and 0 secondary vehicles, or
59 primary and 3 secondary vehicles, or 58 primary and 6 secondary vehicles,
and so on. For a given resource level we run the optimisation algorithm allowing
for one primary to be swapped for three secondary vehicles, and vice versa,
and we report KPIs for the best performing, in terms of maximal
survival, combinations, as well as for the case when the number of secondary
vehicles is fixed at zero. We call these the multiple vehicle type and single
vehicle type scenarios.

Figures~\ref{fig:results_objective_single}-\ref{fig:results_abandoned_multiple}
display the obtained KPIs for each of these allocations. Confidence intervals
over the eight replications were too small to display on the plots. It can
immediately be seen that increasing the resource level has a positive effect on
all KPIs, with vehicle utilisations, mean response times, and percentage
abandoned decreasing, and overall survival increasing. Similarly, as expected
increasing the demand of emergency calls has a negative impact on all KPIs.

\begin{figure}[htb!]
\begin{center}
\begin{subfigure}{0.4\textwidth}
\includegraphics[width=\textwidth]{img/plots/single_Objective}
\caption{Single vehicle type scenario.}
\label{fig:results_objective_single}
\end{subfigure}
\begin{subfigure}{0.4\textwidth}
\includegraphics[width=\textwidth]{img/plots/multiple_Objective}
\caption{Multiple vehicle type scenario.}
\label{fig:results_objective_multiple}
\end{subfigure}
\end{center}
\caption{MESLMHPHF objective expected survival results. \textcolor{red}{(Note that Figures~\ref{fig:results_objective_single} and~\ref{fig:results_objective_multiple} do not have the same y-axis scale in order to keep the data points visible.)}}
\label{fig:results_objective}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\begin{subfigure}{0.4\textwidth}
\includegraphics[width=\textwidth]{img/plots/single_A1Survival}
\caption{Single vehicle type scenario.}
\label{fig:results_a1survival_single}
\end{subfigure}
\begin{subfigure}{0.4\textwidth}
\includegraphics[width=\textwidth]{img/plots/multiple_A1Survival}
\caption{Multiple vehicle type scenario.}
\label{fig:results_a1survival_multiple}
\end{subfigure}
\end{center}
\caption{Expected survival of A1 patients results.}
\label{fig:results_a1survival}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\begin{subfigure}{0.4\textwidth}
\includegraphics[width=\textwidth]{img/plots/single_AmbulanceUtilisation}
\caption{Single vehicle type scenario.}
\label{fig:results_ambulance_utilisation_single}
\end{subfigure}
\begin{subfigure}{0.4\textwidth}
\includegraphics[width=\textwidth]{img/plots/multiple_AmbulanceUtilisation}
\caption{Multiple vehicle type scenario.}
\label{fig:results_ambulance_utilisation_multiple}
\end{subfigure}
\end{center}
\caption{Ambulance utilisation results.}
\label{fig:results_ambulance_utilisation}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[width=0.4\textwidth]{img/plots/multiple_RRVUtilisation}
\end{center}
\caption{RRV utilisation results.}
\label{fig:results_rrv_utilisation_multiple}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\begin{subfigure}{0.4\textwidth}
\includegraphics[width=\textwidth]{img/plots/single_MeanResponseTime}
\caption{Single vehicle type scenario.}
\label{fig:results_response_single}
\end{subfigure}
\begin{subfigure}{0.4\textwidth}
\includegraphics[width=\textwidth]{img/plots/multiple_MeanResponseTime}
\caption{Multiple vehicle type scenario.}
\label{fig:results_response_multiple}
\end{subfigure}
\end{center}
\caption{Mean response time results.}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\begin{subfigure}{0.4\textwidth}
\includegraphics[width=\textwidth]{img/plots/single_PercentAbandoned}
\caption{Single vehicle type scenario.}
\label{fig:results_abandoned_single}
\end{subfigure}
\begin{subfigure}{0.4\textwidth}
\includegraphics[width=\textwidth]{img/plots/multiple_PercentAbandoned}
\caption{Multiple vehicle type scenario.}
\label{fig:results_abandoned_multiple}
\end{subfigure} 
\end{center}
\caption{Percent of abandoned calls results.}
\end{figure}

It is particularly interesting to compare the scenario in which single vehicle
types (only emergency ambulances) were allocated against the scenario where
multiple vehicle types (both emergency ambulances and rapid response vehicles)
were allocated. For equivalent resource levels, introducing secondary vehicles
increases ambulance utilisation, and so decreases availability, resulting in
an increase in abandoned calls. However, introducing secondary vehicles also
results in a decrease in the mean response time, and a large increase in the
expected survival, especially those of speciality A1.
\textcolor{red}{In fact, it can be seen from Figure~\ref{fig:results_a1survival_multiple} that when using multiple vehicle types the survival of A1 patients plateaus as the resource level increases, indicating that the system has reached an upper bound on the survival of A1 patients. This is corroborated by considering the survival function of A1 patients given in Equation~\ref{eqn:survival_curve}, where even if all patients were seen immediately a survival of 43.5\% would be expected, and so accounting for even a minimum travel time to the patients an observed plateau value of 32\% could not be improved upon by adding more resource level.}

An interesting phenomenon occurs when looking at
Figure~\ref{fig:results_response_multiple}. Here it seems that, when using
multiple vehicles, mean response time decreases as demand increases, which is
counter-intuitive. This is however due to a peculiar interplay between demand
and vehicle behaviours: as stated in
Section~\ref{sec:demand_scenarios} only demand for emergency patients are
increased (specialities A1 and A2), which are also the patients that can be seen
by secondary vehicles, and so would have a lower response time than
non-emergency patients (speciality B). Thus, demand scenario \textbf{D45}
receives a larger proportion emergency patients than scenario \textbf{D13}, and
so a larger proportion of patients with lower response times due to being seen
by secondary vehicles; bringing the overall mean response time down. This does
not mean however that emergency patients in scenario \textbf{D45} have lower
response times than those in scenario \textbf{D13}.

It is also noticeable that the increases or decreases in the simulation
derived KPIs are not necessarily monotonic with increases in resource level.
This is due to the derived allocations being based on the MESLMHPHF score
only, which is a measure of survival, while the simulation derived KPIs are
indirect measures of performance. Again, this may indicate that these indirect
measures of performance, such as mean response times or percentage within
target, are not good indicators of survival.

Crucially, the plots show that the allocations produced by the optimisation
algorithm perform better than the current allocation for the same demand and
resource levels. The current levels include 81 EAs and 13 RRVs, so a resource
level of 85.33, with multiple vehicles. Comparing the derived allocations for
this level in
Figures~\ref{fig:results_objective_single}-\ref{fig:results_abandoned_multiple}
to the results given in Table~\ref{tbl:current_grid_results}, the derived
allocations give lower ambulance and RRV utilisations, lower mean response
times, and a higher percentage of patients seen within target.

\section{Discussion \& Conclusions}\label{sec:discussion}
This paper has described the development of EMS demand and capacity models and
their application to the city of Jakarta, Indonesia. To our knowledge, this is
the first such study of analysing emergency ambulance services in Jakarta, and
the work has directly assisted and guided providers and the government to make
investment decisions for a coordinated and free to use EMS.

Firstly, we have considered patient survival and outcomes within a developed
MESLMHPHF model. By considering geospatial travel times and accounting for the
varying needs of different patient types through categorising patients by
speciality and through using different survival function profiles, an expected
overall survival was given. This was used within a optimisation processes, with
an evolutionary metaheuristic, to find vehicle fleet allocations that maximised
expected survival. This expected survival function extends previous work in
\citet{Knight2012918} to include more than one vehicle type. A particular novelty
here was numerically solving implicit utilisation relationships,
equations~\ref{eqn:utilisations_primary} and~\ref{eqn:utilisations_secondary},
to calculate the expected survival, as an alternative to busy fractions and
hypercube models.

Secondly, a discrete event simulation model has been used to evaluate existing
and potential heterogeneous vehicle ambulance feet allocations in terms of key
performance measures, such as response times, survival, and vehicle utilisation
rates. This takes into consideration geospatial demand and travel, and temporal
variation in demand and traffic levels. A novel feature of our approach is that
the model comprises of sequential simulations, feeding data directly from one
into the other in order to simulate primary and secondary vehicles separately
while maintaining synchronicity, with the overall aim of reducing the complexity
of the simulation logic and maintaining modularity.

Both models investigate ambulance service activity in the case of heterogeneous
patients and heterogeneous fleets. Heterogeneous patient groups consider those
with distinct demand profiles, priorities, and survival functions. Heterogeneous
fleets concern different types of vehicles, in our case emergency ambulances
(EAs) which respond to every patient, and Rapid Response Vehicles (RRVs) which
can be utilised to reach patients faster, despite being unable to transport
patients themselves.

Using a combination of models has permitted an approach that can capture
performance measures or take into account factors that each model in isolation
can not. For example, the maximum expected survival model is a lot more
appropriate for use within an optimisation process as the runtimes are a far
more reasonable than the simulation. On the other hand, the simulation allows
for greater complexity such as temporal demand, and is able to capture a
greater range of KPIs. Crucially, using mixed methodology such as this allowed
for independent evaluation of the output of one using the other; but also
insights from an initial run of the simulation gave helped parametrise and
justify assumptions of the optimisation.

A key feature of both models, and their novelty, is the consideration of
heterogeneous fleets. In a highly populated area such as Jakarta (around
16,000 population per km$^2$), the inclusion of RRVs such as paramedics on
motorbikes is very important, given RRVs can access areas that cannot easily
or quickly be accessed by ambulances. Results from both the optimisation and
simulation research showed that the RRVs can help increase the overall
survival and reduce the response times, especially crucial when responding to
life threatening events such as cardiac arrests \citep{holmen2020shortening}.
\textcolor{red}{In building the models, some assumptions or relaxations were
made to simplify the model, and future work should investigate the effects of
these, for example the assumption that service times are not location dependent
is unrealistic, but relaxed here. Further developments of the model could
consider explicitly modelling service rates using the travel time matrices.}

The results of our research also suggest that allocation strategies may not be
intuitive to ambulance service managers, further emphasising the value of a
modelling approach. For example, senior managers suggested the use of a grid
allocation to maximise geographic coverage, with ambulances equally spread
across the city. This was shown to be sub-optimal because neighbourhoods have
different population densities and characteristics. For example, the number of
daily commuters in Jakarta is considerably high during day time due to work
and education \citep{BPS_Jakarta_migrasi}. Municipalities where trading,
educational institutions, and offices are concentrated may become more
populated during day time. The model has quantified the deterioration in
response times and patient outcomes should a fixed grid system be implemented,
leading to 119 to drop this consideration.

The ambulance posts in Jakarta depend on the service provider. Those
ambulances provided by the government are in various locations including
government buildings as well as community clinics and sub-district hospitals.
In our case, we used current ambulance posts as the locations for allocating
the resources. As demands for ambulances may change from time to time, future
work could evaluate different, and perhaps dynamic, ambulance posts that
depend on changing demand volumes in neighbourhoods by time of the day.

The data used for ambulance demand covered one year from 1 January to 31
December 2019, prior the COVID-19 pandemic. Recent studies related to
ambulance demand indicate that the  pandemic has severely impacted on the
utilisation of EMS, for examples in call volumes \citep{csan2021effects} and in
specific medical conditions such as trauma \citep{azbel2021effects}, and
possibly still continue to affect demand. Ambulance providers, such as 118 and
119, may use our modelling tools to incorporate future demand data and
re-evaluate resource needs. To aid this, we are currently working on
interfacing the simulation and optimisation into a single easy to use decision
support tool with a dashboard (data visualiser).

Studies have shown that the quality in pre-hospital data collection varies
considerably in Indonesia \citep{hooper2019_datacollection}. We encountered
similar challenges and have suggested to 119 and the Indonesian Government
that they should continue to strive to improve the coverage and quality of
data collection. Future work could explore in more depth unmet demand that is
not recorded in the ED or in the ambulance data.

Indonesia is a vast country with an uneven population density and varying
quality of available healthcare resources. The models have been parameterised
with data from a high population, high density urban area which is also
relatively well resourced with healthcare facilities compared to other regions
in the country. Ambulance services in rural areas of Indonesia may not have
the same quality in management and organisation compared to the capital,
Jakarta. We intend to conduct future studies to apply the developed models to
analyse ambulance demand and allocations in different areas of Indonesia,
including in rural regions.

In conclusion, the developed models have demonstrated a novel approach in
modelling ambulance allocations that incorporate multiple vehicle types and
health conditions, whilst also capturing patient survival. Our work has
already informed major decisions on the design of a free to use and
coordinated EMS system for Jakarta. Ongoing collaboration will continue to
assist ambulance providers and the Indonesian Government in providing
evidence-based decision making for the benefit of patients and the population
they serve, including the exploration of the roll-out of 119 beyond Jakarta to
other regions of Indonesia.

\section*{Funding}
This work was supported by the Engineering \& Physical Sciences Research Council
(EPSRC) [grant number EP/T003197/1].

\section*{Acknowledgements}
We would like to express our sincere gratitude to Indonesian emergency ambulance
providers including 118 and 119 for their support, insights and provision of
data. In particular we wish to thank Professor Aryono Djuned Pusponegoro and Ms
Asti Puspita Rini, Founder and Director of 118 Ambulance Service Foundation and
Dr Winarto, Head of the 119 Ambulance Service in Jakarta.

\newpage

\appendix

\section{Annotated Objective Function}\label{apx:annotated}
\includegraphics[width=\textwidth]{img/annotate}

\newpage

\textcolor{red}{
\section{Full MESLMHPHF Optimisation Model}\label{apx:fullmodel}
Decision variables: $Z_a$, $\tilde{Z}_a$,
Maximise:
\begin{equation*}
g\left(Z_a, \tilde{Z}_a\right) =
\sum_{p \in \mathcal{P}} \sum_{a \in \mathcal{A}}
\left( \sum_{k \in \mathcal{K}_A}  w_k \lambda_{pk} \hat{\Psi}_{kpa} +
\sum_{k \in \mathcal{K}_B}  w_k \lambda_{pk} \Psi_{kpa} \right)
\end{equation*}
where
\begin{equation*}
\Psi_{kpa} = s_k\left( b_{pa} \right)
\left(1 - \pi_{a}^{Z_a} \right)
\prod_{\alpha \in \mathcal{A}}
\pi_{\alpha}^{\left(Z_{\alpha} \beta_{p\alpha a} \right)}
\end{equation*}
and
\begin{align*}
\hat{\Psi}_{kpa} &= s_k\left(\tilde{b}_{pa}\right)
\left(1 - \tilde{\pi}_{a}^{\tilde{Z}_a} \right)
\prod_{\alpha \in \mathcal{A}}
\tilde{\pi}_{\alpha}^{\left(\tilde{Z}_{\alpha} \beta_{p\alpha a}\right)}
\pi_{\alpha}^{\left(Z_{\alpha} R_{p \alpha a}\right) } \nonumber \\
&+ s_k\left(b_{pa}\right) \left(1 - \pi_{a}^{Z_a} \right)
\prod_{\alpha \in \mathcal{A}}
\pi_{\alpha}^{\left(Z_{\alpha}\beta_{p\alpha a}\right)}
\tilde{\pi}_{\alpha}^{\left(\tilde{Z}_{\alpha}
\left(1 - R _{p a\alpha}\right)\right)}
\end{align*}
subject to
\begin{equation*}
Z_a, \tilde{Z}_a \in \mathbb{N_0}
\end{equation*}
\begin{equation*}
Z_a + \frac{1}{3} \tilde{Z}_a = \mathcal{L}
\end{equation*}
where $\mathcal{L}$ is the resource level.
}

\section{Model Parameters}\label{apx:parameters}
Call arrival rates $\lambda_{pk}$ are derived from 2019 demand data split by
municipality and speciality, and time of day. Probabilities $q_{pky}$ are
similarly derived from the 2019 data of transit journeys. All traffic-free
travel times are found using Google Maps API, while time-dependent traffic
delays are found from the TomTom website and given in
Table~\ref{tbl:delay_factors}.

\begin{figure}[b]
\centering
  \includegraphics[width=0.45\textwidth]{img/time_on_site_fit.pdf}
    \caption{Comparison between the sampled time on site and the lognormal fit.}
  \label{fig:lognorm_fit}
\end{figure}

\begin{table}[b]
\begin{tabular}{ccccc}
\toprule
$h$ & 0000-0500 & 0500-1500 & 1500-1800 & 1800-0000 \\
\midrule
$d_h$ & 0.98 & 0.66 & 0.59 & 0.77 \\
$\tilde{d}_h$ & 1.96 & 0.91 & 0.83 & 1.16 \\
\bottomrule
\end{tabular}
\caption{Primary and secondary delay factors.}
\label{tbl:delay_factors}
\end{table}

The other models are parameterised in the following way:
\begin{itemize}
  \item From a sample of calls the time at site $G_k$ was found to follow a
        lognormal distribution with parameters $\mu=-0.6219$ and $\sigma=0.8048$.
        For the case of Jakarta this was modelled identically for all
        specialities $k$. Comparison between the lognormal fit and the sampled
        times are given in Figure~\ref{fig:lognorm_fit}.
  \item From discussions with staff at the ambulance service in Jakarta, the
        time at hospital $J_k$ was modelled as a Uniform distribution between
        40 and 60 minutes for the emergency specialities A1 and A2, and
        between 20 and 30 minutes for non-emergency speciality B.
  \item From discussions with staff at the ambulance service in Jakarta, the
        refill time $\Theta$ is taken to be 60 minutes for an emergency
        ambulance, and 15 minutes for an RRV.
\end{itemize}

\section{Exploration of Optimisation Hyperparameters}\label{apx:hyperparameters}
Figure~\ref{fig:hyperparameters_exploration} shows the performance of the
evolutionary metaheuristic algorithm under different values of the
hyperparameters $N$, $\kappa$, $m_0$, and $c$.

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{img/hyperparameter_exploration}
\caption{Comparison between optimisation performance for low, medium, and high
values of $N$, $\kappa$, $m_0$, and $c$. Optimisation is run on demand
scenario \textbf{D19}, for a resource level of 75.}
\label{fig:hyperparameters_exploration}
\end{figure}

\section{Improved Allocations for Current Vehicle Numbers}
Figures~\ref{fig:optimal_current_allocation_13},
\ref{fig:optimal_current_allocation_19}, \ref{fig:optimal_current_allocation_34}
and~\ref{fig:optimal_current_allocation_45} show the improved allocations for
81 primary and 13 secondary vehicles under demand scenarios \textbf{D13},
\textbf{D19}, \textbf{D34}, and \textbf{D45} respectively.

\begin{figure}
\begin{center}
\includegraphics[width=0.65\textwidth]{img/map_optimised_13}
\caption{Improved allocation of 81 EAs and 13 RRVs, under scenario \textbf{D13}.}
\label{fig:optimal_current_allocation_13}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.65\textwidth]{img/map_optimised_19}
\caption{Improved allocation of 81 EAs and 13 RRVs, under scenario \textbf{D19}.}
\label{fig:optimal_current_allocation_19}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.65\textwidth]{img/map_optimised_34}
\caption{Improved allocation of 81 EAs and 13 RRVs, under scenario \textbf{D34}.}
\label{fig:optimal_current_allocation_34}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.65\textwidth]{img/map_optimised_45}
\caption{Improved allocation of 81 EAs and 13 RRVs, under scenario \textbf{D45}.}
\label{fig:optimal_current_allocation_45}
\end{center}
\end{figure}

\newpage

% \bibliographystyle{plainnat}
% \bibliographystyle{abbrvnat}
\bibliographystyle{apalike}
% AUTHOR: Include your bib file here
\bibliography{literature}

% %\section*{AUTHOR BIOGRAPHIES}
% \section*{CRediT authorship contribution statement}
% \begin{itemize}
%   \item {\bf Geraint Palmer:} Conceptualisation, Investigation, Methodology, Software, Writing – original draft
%   \item {\bf Mark Tuson:} Conceptualisation, Investigation, Methodology, Software, Writing – original draft
%   \item {\bf Vince Knight:} Conceptualisation, Funding acquisition, Investigation, Methodology, Software, review \& editing.
%   \item {\bf Paul Harper:} Conceptualisation, Investigation, Methodology, Funding acquisition, Resources, Project administration, Writing – review \& editing.
%   \item {\bf Sarie Brice:} Conceptualisation, Investigation, Methodology, Writing – original draft
%   \item {\bf Leanne Smith:} Methodology.
%   \item {\bf Daniel Gartner:} Conceptualisation, Funding acquisition, Methodology, review \& editing.
% \end{itemize}

\end{document}

### 2024-07-26 response/response.tex
address some comments by reviewer 2

@@
 \end{quote}
+This is an interesting suggestion.
+Currently we model the survival function to be independent of the vehicle type, that is $s_k$ is the same regardless if the vehicle that arrived is a primary or secondary vehicle.
+As you note, this may not be realistic, and we may want to be able to model situations where there are different survival functions for each vehicle type.
+However, this becomes more complicated when both vehicles arrive, but at different times - there will be an interaction between the functions that requires more investigation and is beyond the scope of this study.
+These scenarios are interesting, and we have added a sentence in the paper explaining this as further work.
 
@@
 \end{quote}
+I believe the fix should be ``Note that having .... is possible''.
 
@@
 \end{quote}
+Fixed.
 
@@
 \end{quote}
+Fixed throughout.
 
@@
 \end{quote}
+Section 4.4 explains this, it is not an optimisation problem.
+We solve equations 13 and 14, that is we find values of $\lambda_a$ and $\tilde{\lambda}_a$ such that those equations are satisfied.
+These equations are non-linear and so the values are found using the MINPACK hybrd and hybrj algorithms, implemented in Scipy.
 

### 2024-07-26 tex/main.tex
address some comments by reviewer 2

@@
         pick-up location $p$ is taken to hospital $y$.
-        Note that $\sum_{y \in \mathcal{Y}} q_{pky} < 1$ is possible, that is a
+        Note that \textcolor{red}{having} $\sum_{y \in \mathcal{Y}} q_{pky} < 1$ is possible, that is a
         patient may not go to any hospital, in which case the ambulance returns
@@
 hospital ($C_{py}$), spend some time handing over the patient at the hospital
-($J_k$), travel back to it's original location ($D_{ya}$), then spend time
+($J_k$), travel back to its original location ($D_{ya}$), then spend time
 refilling and refuelling ($\Theta$).
@@
 \textcolor{red}{The full MESLMHPHF optimisation model is given in
-Appendix~\ref{apx:fullmodel}, and its components are described fully here.}
+APPENDIX~\ref{apx:fullmodel}, and its components are described fully here.}
 The MESLMHPHF is given by Equation~\ref{eqn:meslmhphf}.
@@
 busy. These interpretations are given as annotations to the equations
-in~\ref{apx:annotated}.
+in~\textcolor{red}{APPENDIX}\ref{apx:annotated}.
+
+\textcolor{red}{Note that in Equations~\ref{eqn:survival_A}
+and~\ref{eqn:survival_B}, the survival functions used, $s_k(t)$, are the same
+whether a primary or a secondary vehicle arrives first. This may be unrealistic,
+where primary vehicles may be better equipped than secondary vehicles to deal
+with some emergencies. One option to overcome this would be to have distinct
+survival functions for each vehicle type, however there may be some interactions
+between the survival probability functions that would not be captured, for
+example when a primary vehicle arrives after a secondary vehicle, there may
+still be a better chance of survival that the secondary vehicle only. This would
+be an interesting avenue of further research. For this work, survival functions
+will be the same for all vehicles, representing the probability of surviving
+given time to be seen by any emergency vehicle.}
 
@@
 preferences $\beta_{p a_1 a_2}$ and $R_{p a_1 a_2}$ are themselves defined by
-travel times (Equations~\ref{eqn:beta} and~\ref{eqn:R}), this may not have much
-affect for this study, \textcolor{red}{as the model already gives preference to
+travel times (Equations~\ref{eqn:beta} and~\ref{eqn:R}), this \textcolor{red}{might not have a strong impact on the results of \st{may not have much
+affect for}} this study, \textcolor{red}{as the model already gives preference to
 the locations with the shortest service times.}
@@
 components. Primary vehicles operate independently of secondary vehicles, that
-is the way primary vehicles respond to a patient is unaffected by the presence
+is\textcolor{red}{,} the way primary vehicles respond to a patient is unaffected by the presence
 or lack of a secondary vehicle.
@@
 simulated after the simulation of primary vehicles, taking as inputs the exact
-list of events that occurred. That is the logic of the secondary vehicles is
+list of events that occurred. That is\textcolor{red}{,} the logic of the secondary vehicles is
 determined by observing the actions of primary vehicles and reacting to them.
@@
 of a typical period for forecasting future needs, hence the decision to use
-the available 2019 demand. \ref{apx:parameters} provides further details on the
+the available 2019 demand. \textcolor{red}{APPENDIX}~\ref{apx:parameters} provides further details on the
 parameterisation of the model, including how travel time estimates were
@@
 average service times of primary and secondary vehicles derived from the initial
-simulation model. \ref{apx:hyperparameters} shows some initial explorations on
+simulation model. \textcolor{red}{APPENDIX}~\ref{apx:hyperparameters} shows some initial explorations on
 the hyper-parameter choices, giving confidence that the chosen parameters, with

### 2024-07-29 response/response.tex
address MINPACK comment, add daniel's revision, fix min/max comment in algorithm

@@
 \end{quote}
+We have now included a paragraph on this in the introduction.
 
@@
 \end{quote}
-The algorithm is correct.
-Line 5 of the algorithm states that $m_0 \leftarrow \lceil m_0 - \max(m_0, i \times c) \rceil$.
-Without the ceiling function, this would have the effect of linearly decreasing wiht $i$ until 0, and then remaining at 0 thereafter.
-With the ceiling function, the linear part becomes a stepwise decent to 0, and then 0 thereafter.
+Thank you, this has now been fixed.
 
@@
 \end{quote}
+We have now added this.
 
@@
 \end{quote}
-Section 4.4 explains this, it is not an optimisation problem.
-We solve equations 13 and 14, that is we find values of $\lambda_a$ and $\tilde{\lambda}_a$ such that those equations are satisfied.
-These equations are non-linear and so the values are found using the MINPACK hybrd and hybrj algorithms, implemented in Scipy.
+We have added further explanation detailing how we solve the system of non-linear equations, and have added a reference to the underlying optimisation algorithm used.
 

### 2024-07-29 tex/main.tex
address MINPACK comment, add daniel's revision, fix min/max comment in algorithm

@@
 could be readily applied to other locations.
+
+\textcolor{red}{There are several managerial implications for EMS providers and
+government agencies in Jakarta and Indonesia, respectively. The findings
+emphasise the importance of adopting a dynamic and data-driven approach to
+resource allocation. The inclusion of a heterogeneous fleet, such as Rapid
+Response Vehicles (RRVs) alongside traditional ambulances, has been shown to
+significantly improve response times and patient survival rates, particularly in
+densely populated urban areas. Managers are advised to invest in these diverse
+fleets and adopt flexible allocation strategies that account for temporal
+variations in demand, such as the daily influx of commuters and the differing
+population densities across neighbourhoods. The research underscores the
+necessity of enhancing data collection practices to support informed
+decision-making, with the development of user-friendly decision support tools to
+assist in visualising data and predicting outcomes}
 
@@
 hybrj algorithms by using Scipy's \texttt{fsolve} function
-\citep{2020SciPy-NMeth}. It should be noted that in this section we assume that
+\citep{2020SciPy-NMeth}.
+\textcolor{red}{These algorithms are efficient methods for finding zeros of systems of non-linear equations.
+They are based on the Powell hybrid method \cite{powell70}.} It should be noted that in this section we assume that
 all $\lambda_{pk}$ are static and not time-dependant, which is an assumption
@@
   \State Keep the top $\kappa$ solutions\;
-  \State $m \leftarrow \lceil m_0 - \max\left(m_0, i \times c\right)\rceil$\;
+  \State $m \leftarrow \lceil m_0 - \min\left(m_0, i \times c\right)\rceil$\;
   \For{$j \leftarrow 0$ \algorithmicto $N - \kappa$}

### 2024-07-30 response/response.tex
address comments on using simulation survival, adding variability measures

@@
 \end{quote}
+Thank you for the comment, standard deviations of mean response times are now given in Table 1 and 4.
+This is a particularly useful inclusion as the large variance in the response times explains the discrepencies in the simulated and expected survivals.
 
@@
 \end{quote}
+We have now added new subsection comparing survival obtained through the MESLMHPHF and through the simulation.
+There are discrepancies between the two, and in the new subsection we discuss reasons for this, and offer some directions of future work that could help in alleviating some of these problems.
 
@@
 \end{quote}
+This is a very good point, and is similar to the situation when a primary vehicle reaches the patient before the secondary vehicle.
+Ideally in these cases the later vehicle would turn around and be free to pick up another patient.
+However one assumption of the model is that en-route communications are difficult and do not occur.
+We have now emphasised this in the text.
 
@@
 \end{quote}
+We did not consider using the simulation as part of the optimisation process, and is beyong the scope of the paper.
+However, we think this is a great idea, and have discussed this as future work in the paper, as a tool to give greater consideration to the variability in the system.
 

### 2024-07-30 tex/main.tex
address comments on using simulation survival, adding variability measures

@@
 However, hard cut-off curves like Equation~\ref{eqn:survival_cutoff}, with a
-cutoff of $L$ can still be used to represent meeting artificially selected
+cut-off of $L$ can still be used to represent meeting artificially selected
 targets, for example for transportation jobs.
@@
 \caption{Survival function $s(t)$, estimated by \citet{Valenzuela20001206}
-compared with $s_L(t)$ with a hard cutoff of $L=8$, represented as a step
+compared with $s_L(t)$ with a hard cut-off of $L=8$, represented as a step
 function.}
@@
   still reach the pickup locations, due to potential difficulties in
-  communication en-route.
+  communication en-route. \textcolor{red}{Consider also the situation when a
+  secondary vehicle arrives and the patient does not require a hospital, with
+  better en-route communication this would mean the primary vehicle turns
+  around, however without en-route communication which is assumed here, the
+  primary vehicle must still reach the patient pick-up location.}
 \end{itemize}
@@
 Mean Response Time (mins) & 17.69 & 22.29 & 24.05 \\
+\textcolor{red}{StDev Response Time (mins)} & \textcolor{red}{6.98} & \textcolor{red}{8.02} & \textcolor{red}{7.94} \\
 MESLMHPHF Objective & 98.34\% & 87.95\% & 85.64\% \\
@@
 response time, due to their survival function Equation~\ref{eqn:survival_curve}.
+\textcolor{red}{There is a large discrepancy between the expected survival
+measures calculated from the MESLMHPHF objective, and the observed survival from
+the simulation. This is discussed further in the next section.}
 It is worth noting that for the simulation based KPIs (mean response times and
@@
 \end{table}
+
+\textcolor{red}{
+\subsection{Comparing Expected \& Simulated Survival Measures}
+In Table~\ref{tbl:current_optimal_compare} it can be seen that there is a large
+discrepancy between the expected survival, both overall and of A1 patients,
+computed using the MESLMHPHF described in Section~\ref{sec:meslmhphf} and
+calculated using the simulation. This is a consequence of the variability in
+response times, which is explicitly modelled by the simulation but not captured
+by the MESLMHPHF. As seen in the Table, the standard deviations of the response
+times are relatively large in comparison to the mean response times, and so this
+effect would be substantial here.}
+
+\textcolor{red}{
+The discrepancies are exacerbated by the use of cut-off survival function curves
+for most of the patients: consider a survival function with cut-off $L$, with
+response time below $L$ the patient's chance of survival is 0, and above $L$ the
+patient's chance of survival is guaranteed. The optimisation will find a
+solution such that will ensure that the expected response time is above $L$,
+possibly as low as $L + \epsilon$, but will not attempt to push the expected
+response time above this, as it will have no effect on the expected survival.
+However considering variability of mean response times, a considerable
+proportion of the response times will lie below the expected value, and so
+possibly below $L$, resulting in worse survival probabilities, as observed in
+Table~\ref{tbl:current_optimal_compare}. Therefore the distributions chosen to
+model travel times can have a large impact on the survival measures, in this
+case the naive Triangular distribution.}
+
+\textcolor{red}{
+This leads to a number of avenues of future work to improve this and overcome
+this effect. Variability could be explicitly modelled within the MESLMHPHF, such
+that it is not only maximising \textit{expected} survival, but considers some
+percentiles of guaranteed survival. Another avenue would be to use the
+simulation explicitly in the optimisation process, incorporating the simulation
+into the objective function and using simulation-optimisation techniques to
+improve runtime efficiency. However, as discussed in Section~\ref{sec:survival},
+maximising expected survival of all patients can lead to a fairer allocation of
+resources, with difficult to reach patients being given equal consideration to
+closer, easier to reach patients. When considering the variability of travel
+times in the optimisation process, difficult to reach patients may be ignored, indistinguishable from patients who happen to sample a response time at the
+extreme tail of a given distribution curve.}
+
+\textcolor{red}{
+Therefore, due to this interpretation of fairness of allocations, and due to the
+naive choice of travel time distributions, including difficulties in being able
+to validate such a model and the choices made, we consider only the MESLMHPHF
+objective from now on.
+}
 

### 2024-07-31 tex/main.tex
add consideration of coverage and response time objectives

@@
 
+\textcolor{red}{Finding allocations that maximise survival can result in
+allocations with different behaviours to those found by minimising response
+times or maximising coverage. Coverage maximisation models are often easier to
+solve, and easier to communicate, but may not take into consideration severity
+or intensity of demand \citep{Erkut200842}. This results in a minority of
+patients, particularly those more isolated, being over-serviced, that is, having
+more relative resources allocated to their needs that the majority, at the
+expense of the majority. Minimising average response times on the other hand
+treats all patients equally, but may have the unintended outcome of disregarding
+some more further out patients whose response times would be larger that others
+regardless of the allocation. It would still disregard the severity or type of
+the patient. Allocations based on maximising survival are similar to those that
+minimise response time, overcoming the issue of over-servicing outlying
+patients, though it is still possible to disregard them completely. It has an
+advantage over minimising response times as it can account for patient severity.
+Having survival as functions of response times has the effect of weighting gains
+in response times by how valuable those gains are to a patient.}
+
 \subsection{Maximal Expected Survival Location Model with Heterogeneous
@@
 will be the same for all vehicles, representing the probability of surviving
-given time to be seen by any emergency vehicle.}
+given time to be seen by any emergency vehicle.
+Note also that the survival functions are functions of $b_{pa}$, the expected
+travel time between $p$ and $a$, and do not consider the variability in the
+travel times. This will be discussed further alongside results in
+Section~\ref{sec:compare_exp_and_sim}.}
 
@@
 \textcolor{red}{These algorithms are efficient methods for finding zeros of systems of non-linear equations.
-They are based on the Powell hybrid method \cite{powell70}.} It should be noted that in this section we assume that
+They are based on the Powell hybrid method \citep{powell70}.} It should be noted that in this section we assume that
 all $\lambda_{pk}$ are static and not time-dependant, which is an assumption
@@
 
-\subsection{Planning for Future Demand}\label{sec:demand_scenarios}
+\subsection{Demand Scenarios}\label{sec:demand_scenarios}
 Similarly to many other cities worldwide, ambulance services in Jakarta
@@
 \textcolor{red}{
-\subsection{Comparing Expected \& Simulated Survival Measures}
+\subsection{Comparing Expected \& Simulated Survival Measures}\label{sec:compare_exp_and_sim}
 In Table~\ref{tbl:current_optimal_compare} it can be seen that there is a large
@@
 response times, which is explicitly modelled by the simulation but not captured
-by the MESLMHPHF. As seen in the Table, the standard deviations of the response
+by the MESLMHPHF. As seen in the table, the standard deviations of the response
 times are relatively large in comparison to the mean response times, and so this
@@
 patient's chance of survival is guaranteed. The optimisation will find a
-solution such that will ensure that the expected response time is above $L$,
-possibly as low as $L + \epsilon$, but will not attempt to push the expected
+solution aim to push the expected response time above $L$, and once this is
+reached, possibly with an expected response time as low as $L$ itself, it will
+not attempt to push the expected
 response time above this, as it will have no effect on the expected survival.
@@
 into the objective function and using simulation-optimisation techniques to
-improve runtime efficiency. However, as discussed in Section~\ref{sec:survival},
-maximising expected survival of all patients can lead to a fairer allocation of
-resources, with difficult to reach patients being given equal consideration to
-closer, easier to reach patients. When considering the variability of travel
-times in the optimisation process, difficult to reach patients may be ignored, indistinguishable from patients who happen to sample a response time at the
-extreme tail of a given distribution curve.}
+improve runtime efficiency.}
 
 \textcolor{red}{
-Therefore, due to this interpretation of fairness of allocations, and due to the
-naive choice of travel time distributions, including difficulties in being able
-to validate such a model and the choices made, we consider only the MESLMHPHF
-objective from now on.
+Due to the naive choice of travel time distributions, including difficulties in
+being able to validate such a model and the choices made, we consider only the
+MESLMHPHF objective from now on.
 }
@@
 life threatening events such as cardiac arrests \citep{holmen2020shortening}.
+
 \textcolor{red}{In building the models, some assumptions or relaxations were
@@
 is unrealistic, but relaxed here. Further developments of the model could
-consider explicitly modelling service rates using the travel time matrices.}
+consider explicitly modelling service rates using the travel time matrices.
+Another limitation identified was the lack of consideration of variability in
+travel times in the optimisation model, which cased some discrepancies between
+the simulated and expected survivals. One method to overcome this could be to
+incorporate variability directly into the MESLMHPHF objective, for example using
+percentiles instead of averages, or to optimise using the simulation as an
+objective. However in both approaches, a good estimate of travel time
+distributions would be needed, with better validation.
+}
 

### 2024-07-31 response/response.tex
add consideration of coverage and response time objectives
Complete first draft of response and review

@@
 \end{quote}
+Done.
 
@@
 \end{quote}
+Thorough verification of the models is done by the automated code testing suits provided in the code base, a paragraph on this is now included under Section 5.5.
+Validation of this model was difficult, and relied on the expert opinion of our collaborators ar 118 and 119.
+This is now discussed in Section 6.
 
@@
 \end{quote}
+A paragraph has now been added to Section 4.1 comparing allocations that maximise survival, maximise coverage, and minimise response times.
 
@@
 \end{quote}
+We did not consider using an exact method to solve this as the were so many non-linear elements to the problem, namely the objective function and the utilisation considerations.
+An exact method would be beyond the scope of this paper, which focusses on the non-linear objective function formulation, and simulation modelling.
 
@@
 \end{quote}
+A reference has now been given justifying the use of triangular distributions when little information is known.
+All other travel time information is given in the Appendix, where we state that travel times are derived using Google Maps API, and traffic delays found using the TomTom API.
 
@@
 \end{quote}
+The call arrival rates are taken directly from provided data, and reported in the Appendix.
+You make a good point regarding the pick-up locations with particularly low arrival rates - this is not a problem for \textbf{D13} as this is what was observed, but might result in artificially low arrival rates when multiplying up by demand.
+This is now discussed in the paper.
 

### 2024-07-31 tex/main.tex
Complete first draft of response and review

@@
 calculated expected travel time, between 75\% and 125\% of the value.
+\textcolor{red}{A Triangular distribution is a useful distribution to use when
+not much is known about the travel times except some basic summary statistics
+\citep{robinson2014simulation}, in this case a likely range of values.}
 
@@
 
-\subsection{Combined Model}
+\subsection{Combined Model}\label{sec:combined_model}
 Combining the logic presented in Sections~\ref{sec:simulation_primary}
@@
 and is described in Section~\ref{sec:survival}.
-The source code for the combined simulation, is available at \citet{zenodo_sim} and
-development took place at \url{https://github.com/MarkTuson/ambulance_simulation}.
+
+\textcolor{red}{
+Verification of a model is ensuring that the conceptual model is implemented
+correctly and without error, while validation of a model ensures that the
+conceptual model accurately represents the real system to be modelled
+\citep{robinson2014simulation}. Verifying the simulation code, alongside
+verifying the optimisation and MESLMHPHF objective function code, is done
+through automated code testing \citep{percival2014test, palmer2023implementing}.
+This includes unit testing to ensure that each sub-component works are requires,
+and is bug and error free; end to end testing, ensuring that the components
+interact together as intended; and extreme value testing ensuring that changes
+in input parameters result in sensible and expected changes in output results
+and behaviours. The testing suits are included in the open code repositories
+provided.
+}
+
+The source code for the combined simulation, is available at \citet{zenodo_sim}
+and development took place at \url{https://github.com/MarkTuson/ambulance_simulation}.
 
@@
 of a typical period for forecasting future needs, hence the decision to use
-the available 2019 demand. \textcolor{red}{APPENDIX}~\ref{apx:parameters} provides further details on the
-parameterisation of the model, including how travel time estimates were
-obtained.
+the available 2019 demand. \textcolor{red}{APPENDIX}~\ref{apx:parameters}
+provides further details on the parameterisation of the model, including how
+travel time estimates were obtained.
+\textcolor{red}{Verification of the models were discussed in
+Section~\ref{sec:combined_model}, however parameter validation for this
+particular case study is difficult, with all primary data available to us being
+used to parametrise the model, still leaving many aspects to be determined by
+expert opinion. This expert opinion however was vital in giving a basic
+validation of the model, with staff from Ambulans 118 and the 119 Emergency
+Ambulance Service in Jakarta approving model results parametrised from their
+provided data and expert opinion.}
 
@@
 
-We consider four different demand scenarios, developed by considering the
+\textcolor{red}{In other cities with more detailed data collection, statistical
+models can be used to predict future spacial demand across a city, as in
+\cite{nicoletta2022bayesian}. In the case of Jakarta, collected call data is
+not detailed enough, but we do}
+consider four different demand scenarios, developed by considering the
 result of our cross-sectional study of patients attending EDs in Jakarta
@@
 remains unchanged.
+\textcolor{red}{Therefore, in each scenario, we multiply the call demand for
+each municipality, $\lambda_{pk}$ by $\frac{13\%}{13\%}=1$,
+$\frac{19\%}{13\%}=1.46$, $\frac{34\%}{13\%}=2.62$, and
+$\frac{45\%}{13\%}=3.46$ respectively, for all $k \in \mathcal{K}_A$,
+$p \in \mathcal{P}$. Note that for some pick-up locations $p$, $\lambda_{pk}$
+is very small, possibly zero, which is realistic in scenario \textbf{D13}, as
+that is what is observed, but may result in artificially low demand in those
+locations when using the multiplication method to scale up demand.}
 

### 2024-08-12 response/response.tex
addresses comments from Vince and Daniel

@@
 \end{quote}
-Fixed, all instances should the $J_k$.
+Fixed, all instances should be $J_k$.
 
@@
 An exact method would be beyond the scope of this paper, which focusses on the non-linear objective function formulation, and simulation modelling.
+We have added using exact methods to the future work discussion.
 
@@
 Thorough verification of the models is done by the automated code testing suits provided in the code base, a paragraph on this is now included under Section 5.5.
-Validation of this model was difficult, and relied on the expert opinion of our collaborators ar 118 and 119.
+We validated the model based on expert opinion from our collaborators in Jakarta at 118 and 119, and several workshops have been carried out in Indonesia where the model was shared.
 This is now discussed in Section 6.

### 2024-08-12 tex/main.tex
addresses comments from Vince and Daniel

@@
   still reach the pickup locations, due to potential difficulties in
-  communication en-route. \textcolor{red}{Consider also the situation when a
+  communication en-route. \textcolor{red}{Consider a situation when a
   secondary vehicle arrives and the patient does not require a hospital, with
@@
 in input parameters result in sensible and expected changes in output results
-and behaviours. The testing suits are included in the open code repositories
+and behaviours. The testing suites are included in the open code repositories
 provided.
@@
 \textcolor{red}{Verification of the models were discussed in
-Section~\ref{sec:combined_model}, however parameter validation for this
-particular case study is difficult, with all primary data available to us being
-used to parametrise the model, still leaving many aspects to be determined by
-expert opinion. This expert opinion however was vital in giving a basic
-validation of the model, with staff from Ambulans 118 and the 119 Emergency
-Ambulance Service in Jakarta approving model results parametrised from their
-provided data and expert opinion.}
+Section~\ref{sec:combined_model}, we validated based on expert opinion from
+our collaborators, Ambulans 118 and the 119 Emergency Ambulance Service in
+Jakarta, where several workshops have been carried out in Indonesia where the
+model was shared and discussed. This expertise was vital in giving a basic
+validation of the model, with collaborators approving model results
+parametrised from their provided data and expert opinion.}
 
@@
 response time, due to their survival function Equation~\ref{eqn:survival_curve}.
-\textcolor{red}{There is a large discrepancy between the expected survival
+\textcolor{red}{There is a discrepancy between the expected survival
 measures calculated from the MESLMHPHF objective, and the observed survival from
@@
 patient's chance of survival is guaranteed. The optimisation will find a
-solution aim to push the expected response time above $L$, and once this is
+solution aiming to push the expected response time above $L$, and once this is
 reached, possibly with an expected response time as low as $L$ itself, it will
@@
 distributions would be needed, with better validation.
-}
+Another direction of future work is to find exact methods of solving the
+MESLMHPHF instead of relying on a metaheuristic algorithm.}
 

### 2024-08-12 tex/main.tex
add comments from Paul

@@
   communication en-route. \textcolor{red}{Consider a situation when a
-  secondary vehicle arrives and the patient does not require a hospital, with
-  better en-route communication this would mean the primary vehicle turns
-  around, however without en-route communication which is assumed here, the
-  primary vehicle must still reach the patient pick-up location.}
+  secondary vehicle arrives and the patient does not require transportation to
+  a hospital, with better en-route communication this would mean the primary
+  vehicle turns around, however without en-route communication which is
+  assumed here, the primary vehicle must still reach the patient pick-up
+  location.}
 \end{itemize}
@@
 Asti Puspita Rini, Founder and Director of 118 Ambulance Service Foundation and
-Dr Winarto, Head of the 119 Ambulance Service in Jakarta.
+Dr Winarto, Executive Director of the 119 Ambulance Service (now PK3D) in Jakarta.
 

### 2025-01-21 response_2/response.tex
Fix column name.

\documentclass{article}

\usepackage{fullpage}
\usepackage{parskip}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{hyperref}

\title{Responses to Reviewers}
\author{}
\date{}

\begin{document}

\maketitle

We would like to express our gratitude to the editor, associate editor, and reviewers for facilitating the review process, and for their useful suggestions and remarks on the manuscript.

This document will address the specific concerns raised.

\section{Responses to the Editor}

There were no comments from the Editor.

\section{Responses to the Associate Editor}
\begin{quote}
``Thank you for your revised manuscript. While one reviewer is satisfied with
    the changes, the second reviewer has raised additional concerns that require
    further clarification. It is clear that you have addressed many of the
    previous remarks; however, some points still need more detailed responses.
    Please provide a clear and thorough explanation for the points raised by the
    second reviewer in your next revision. We look forward to your updated
    manuscript.''
\end{quote}

We are delighted to see that we have addressed all the points raised by reviewer
2. Through this response we will endeavour to demonstrate clearly and precisely
how we have addressed the remaining points raised by reviewer 1.

\section{Responses to the Reviewers}

\subsection{Comments from Reviewer 1}

\begin{itemize}

\item Reviewer 1's comment from the first review:

\begin{quote}
``No measures of variance are given for the simulation results in Tables 1
and 4. Especially since repeated simulation results were conducted, measu-
res of variance should be included to investigate the statistical significance
of the results.''
\end{quote}

\item Our response to reviewer 1's comment from the first review:

Thank you for the comment, standard deviations of mean response times are now
given in Table 1 and 4. This is a particularly useful inclusion as the
large variance in the response times explains the discrepencies in the
simulated and expected survivals.

\item Review 1's new comment:

\begin{quote}
Not addressed: The standard deviation of the response time was inclu-
ded, but no information about the variance of the results over the eight
replications was given. To obtain knowledge about the reliability of the
simulation results, some information should be given about how much the
simulation results differed between the eight replications mentioned on
page 17.
\end{quote}

A short discussion about the variability over the repetitions (including
variance) has now been included. 
% TODO VK: We need to do this, I believe it is only Geraint who has the data.

\item Reviewer 1's comment from the first review:

\begin{quote}
``Unclear, why the subchapter 4.1 on survival functions is necessary, sin-
ce traditional coverage targets are used in the case study and the model
formulation does not depend on characteristics of the survival functions.
While survival functions are an important performance metric in the re-
viewers’ opinion, it is not clear why the reader needs this information to
follow the remainder of the pape.''
\end{quote}

\item Our response to reviewer 1's comment from the first review:

We disagree with the reviewer here. In the case study, traditional targets such
a mean response times are reported alongside survival measures, based on
the survival functions. However, a key part of the paper is the
optimisation process, and all scenarios compared are `optimised'
scenarios that result from this process. The objective function of the
optimisation process maximises survival, based on survival functions.
This is evnidenced in Equations 6 and 7, which are component parts of
the objective function given in Equation 5. Therefore the survival
functions are central to the work presented.

\item Review 1's new comment:

\begin{quote}
''Not addressed: It was expanded upon why survival functions can be sui-
table performance metrics. In the case study however, coverage targets
were used (8 min target for A1 patients, 15 min target for A2 patients
and 60 min target for B patients). It is therefore still unclear, why the
reader needs the information in chapter 4.1. Additionally, there is no dis-
cussion as to why coverage targets were chosen in the case study, after
it was argued, that survival functions can solve some drawbacks of using
coverage targets.``
\end{quote}

and

\begin{quote}
    ``
There is contradicting information about which survival functions were
used in the case study. On page 15 and 16, response time targets for the
different categories are listed. This implies to the reader, that step-wise
survival functions, in other words coverage targets, were used in the case
study. On page 19, line 44 it is stated, that Equation 3 was used as the
survival function for A1 patients. It needs to be clarified, which survival
functions were used in the MESLMHPHF in the case study both for the
optimization metaheuristic and the evaluation of the results.
    ''
\end{quote}

\item Our response:

We have added a sentence to the text to clarify that despite the policy target
of 8 minutes for the A1 patients, we have in fact used the survival function of
equation (3) for A1 patients. This was not explicit and we appreciate the
reviewer insisting this point.

\item Reviewer 1's comment:

\begin{quote}
``The simulated survival probabilities in Table 4 show, that the survival for
A1 patients was not improved in any of the four scenarios in the simu-
lation. As the simulation model is a more realistic model of the actual
1
EMS system in comparison to the MESLMHPHF, the simulated survi-
val probabilities carry more weight in interpreting the results compared
to the expected survival probabilities calculated with the MESLMHPHF.
On page 19, line 42 - 44 however it is stated, that ‘in particular A1 pati-
ents are benefitting from the improved allocations’. It is unclear, how the
authors draw this conclusion, when the simulation results show worse or
equal survival probabilities for A1 patients in all scenarios.''
\end{quote}

The reviewer raises a good point here. We have modified the text to hopefully
clarify that:

\begin{itemize}
    \item The sentence regard A1 patients meant to reflect more on the
        efficiency and accuracy of the optimisation heuristic which we find is
        one of the main contributions of this paper (specifically the novel
        approach used to deal with the circular nature of utilisation and
        allocation).
    \item The optimised allocation give more robust allocations under increasing
        demand although we admit that there is not enough numerical evidence to
        be sure of this effect.
\end{itemize}

Furthermore we have more explicitly referred to our explanation about the cause
of this discrepancy.

Fundamentally, again, the reviewer raises a valid and interesting avenue for investigation and
one that we feel
could not be precisely considered without the work in this paper. We hope to
tackle building a more accurate objective function, that better captures the
stochastic effects, that could be then optimised with our novel algorithm. This,
does not however fit in this paper. If the reviewer would like us to make that
more explicit and clear in the text we would be happy to follow any guidance
they feel is necessary.

\item Reviewer 1's comment:

    \begin{quote}
        ``      Equations 8 and 9: It is stated, that \(\beta_{pa_1a_2}\) indicates, that a vehicle from
        \(a_1\) can reach \(p\) quicker, than a vehicle from \(a_2\). The equations 8 and 9
        however use \(\leq\) and not \(<\) , which is not consistent with the description.
Has been fixed for \(\beta_{pa_1a_2}\) in Equation 8, but not for \(R_{pa_1a_2}\) in Equation
9. ''
    \end{quote}
    This has been fixed.
    % TODO Check with Geraint that this is correct.

\end{itemize}

\subsection{Comments from Reviewer 2}

There were no further comments from Reviewer 2.

\end{document}

### 2025-01-21 tex/main.tex
Fix column name.

@@
 there has been very little prior work on developing an EMS strategy, driven
-by \textcolor{red}{\st{inadequate}} a lack of financial investment and human capital
+by a lack of financial investment and human capital
 \citep{plummer2017ems,pusponegoro2003terrorism,yusvirazi2018state}.
@@
 
-\textcolor{red}{There are several managerial implications for EMS providers and
+There are several managerial implications for EMS providers and
 government agencies in Jakarta and Indonesia, respectively. The findings
@@
 decision-making, with the development of user-friendly decision support tools to
-assist in visualising data and predicting outcomes}
+assist in visualising data and predicting outcomes.
 
@@
         pick-up location $p$ is taken to hospital $y$.
-        Note that \textcolor{red}{having} $\sum_{y \in \mathcal{Y}} q_{pky} < 1$ is possible, that is a
+        Note that having $\sum_{y \in \mathcal{Y}} q_{pky} < 1$ is possible, that is a
         patient may not go to any hospital, in which case the ambulance returns
@@
 
-\textcolor{red}{Finding allocations that maximise survival can result in
+Finding allocations that maximise survival can result in
 allocations with different behaviours to those found by minimising response
@@
 Having survival as functions of response times has the effect of weighting gains
-in response times by how valuable those gains are to a patient.}
+in response times by how valuable those gains are to a patient.
 
@@
 availability where appropriate.
-\textcolor{red}{The full MESLMHPHF optimisation model is given in
-APPENDIX~\ref{apx:fullmodel}, and its components are described fully here.}
+The full MESLMHPHF optimisation model is given in
+APPENDIX~\ref{apx:fullmodel}, and its components are described fully here.
 The MESLMHPHF is given by Equation~\ref{eqn:meslmhphf}.
@@
 
-Here $R_{p a_1 a_2}$ and $\beta_{p a_1 a_2}$ \textcolor{red}{define an ordering
-of the ambulance locations, they} are binary variables indicating the preference
+Here $R_{p a_1 a_2}$ and $\beta_{p a_1 a_2}$ define an ordering
+of the ambulance locations, they are binary variables indicating the preference
 of sending a vehicle from one station to another, that is the dispatch rule.
@@
 at $a_1$ can reach $p$ quicker than a secondary vehicle at $a_2$, defined in
-Equation~\ref{eqn:R}. \textcolor{red}{In order for these to define an ordering
+Equation~\ref{eqn:R}. In order for these to define an ordering
 it is assumed that no two distinct ambulance locations are exactly equidistant
-from a pickup location.} Note that here these parameters are defined using
+from a pickup location. Note that here these parameters are defined using
 travel times, but can be generalised to account for any type of preference,
@@
 busy. These interpretations are given as annotations to the equations
-in~\textcolor{red}{APPENDIX}\ref{apx:annotated}.
-
-\textcolor{red}{Note that in Equations~\ref{eqn:survival_A}
+in APPENDIX~\ref{apx:annotated}.
+
+Note that in Equations~\ref{eqn:survival_A}
 and~\ref{eqn:survival_B}, the survival functions used, $s_k(t)$, are the same
@@
 travel times. This will be discussed further alongside results in
-Section~\ref{sec:compare_exp_and_sim}.}
+Section~\ref{sec:compare_exp_and_sim}.
 
@@
 preferences $\beta_{p a_1 a_2}$ and $R_{p a_1 a_2}$ are themselves defined by
-travel times (Equations~\ref{eqn:beta} and~\ref{eqn:R}), this \textcolor{red}{might not have a strong impact on the results of \st{may not have much
-affect for}} this study, \textcolor{red}{as the model already gives preference to
-the locations with the shortest service times.}
+travel times (Equations~\ref{eqn:beta} and~\ref{eqn:R}), this might not have a strong impact on the results of \st{may not have much
+affect for} this study, as the model already gives preference to
+the locations with the shortest service times.
 
@@
 hybrj algorithms by using Scipy's \texttt{fsolve} function
-\citep{2020SciPy-NMeth}.
-\textcolor{red}{These algorithms are efficient methods for finding zeros of systems of non-linear equations.
-They are based on the Powell hybrid method \citep{powell70}.} It should be noted that in this section we assume that
+\citep{2020SciPy-NMeth}. These algorithms are efficient methods for finding zeros of systems of non-linear equations. 
+They are based on the Powell hybrid method \citep{powell70}. It should be noted that in this section we assume that
 all $\lambda_{pk}$ are static and not time-dependant, which is an assumption
@@
 \For{$i \leftarrow 0$ \algorithmicto $H$}
-  \State Rank the population \textcolor{red}{by decreasing expected survival} according to Equation~\ref{eqn:meslmhphf}\;
+  \State Rank the population by decreasing expected survival according to Equation~\ref{eqn:meslmhphf}\;
   \State Keep the top $\kappa$ solutions\;
@@
 calculated expected travel time, between 75\% and 125\% of the value.
-\textcolor{red}{A Triangular distribution is a useful distribution to use when
+A Triangular distribution is a useful distribution to use when
 not much is known about the travel times except some basic summary statistics
-\citep{robinson2014simulation}, in this case a likely range of values.}
+\citep{robinson2014simulation}, in this case a likely range of values.
 
@@
 components. Primary vehicles operate independently of secondary vehicles, that
-is\textcolor{red}{,} the way primary vehicles respond to a patient is unaffected by the presence
+is, the way primary vehicles respond to a patient is unaffected by the presence
 or lack of a secondary vehicle.
@@
 simulated after the simulation of primary vehicles, taking as inputs the exact
-list of events that occurred. That is\textcolor{red}{,} the logic of the secondary vehicles is
+list of events that occurred. That is, the logic of the secondary vehicles is
 determined by observing the actions of primary vehicles and reacting to them.
@@
   still reach the pickup locations, due to potential difficulties in
-  communication en-route. \textcolor{red}{Consider a situation when a
+  communication en-route. Consider a situation when a
   secondary vehicle arrives and the patient does not require transportation to
@@
   assumed here, the primary vehicle must still reach the patient pick-up
-  location.}
+  location.
 \end{itemize}
@@
 
-\textcolor{red}{
 Verification of a model is ensuring that the conceptual model is implemented
@@
 provided.
-}
 
@@
   time of 8 minutes. In our data set there were 168 identified calls that met
-  this criterion.
+        this criterion. \textcolor{red}{As in~\cite{Knight2012918},
+        equation~(\ref{eqn:survival_curve}) is here used to measure survival of these patients.}
   \item \textbf{A2 - Other emergency patients}: urgent patients that require
@@
 of a typical period for forecasting future needs, hence the decision to use
-the available 2019 demand. \textcolor{red}{APPENDIX}~\ref{apx:parameters}
+the available 2019 demand. APPENDIX~\ref{apx:parameters}
 provides further details on the parameterisation of the model, including how
 travel time estimates were obtained.
-\textcolor{red}{Verification of the models were discussed in
+Verification of the models were discussed in
 Section~\ref{sec:combined_model}, we validated based on expert opinion from
@@
 validation of the model, with collaborators approving model results
-parametrised from their provided data and expert opinion.}
+parametrised from their provided data and expert opinion.
 
@@
 in Table~\ref{tbl:current_grid_results}. The simulation was run over half a
-year, with a month warm up time, over eight replications. We observe that the
+year, with a month warm up time, over eight replications. 
+\textcolor{red}{The variability over the eight replications ...}
+We observe that the
 two proposed grid allocations perform much worse than the current allocation as
@@
 Mean Response Time (mins) & 17.69 & 22.29 & 24.05 \\
-\textcolor{red}{StDev Response Time (mins)} & \textcolor{red}{6.98} & \textcolor{red}{8.02} & \textcolor{red}{7.94} \\
+StDev Response Time (mins) & 6.98 & 8.02 & 7.94 \\
 MESLMHPHF Objective & 98.34\% & 87.95\% & 85.64\% \\
@@
 
-\textcolor{red}{In other cities with more detailed data collection, statistical
+In other cities with more detailed data collection, statistical
 models can be used to predict future spacial demand across a city, as in
 \cite{nicoletta2022bayesian}. In the case of Jakarta, collected call data is
-not detailed enough, but we do}
+not detailed enough, but we do
 consider four different demand scenarios, developed by considering the
@@
 remains unchanged.
-\textcolor{red}{Therefore, in each scenario, we multiply the call demand for
+Therefore, in each scenario, we multiply the call demand for
 each municipality, $\lambda_{pk}$ by $\frac{13\%}{13\%}=1$,
@@
 that is what is observed, but may result in artificially low demand in those
-locations when using the multiplication method to scale up demand.}
+locations when using the multiplication method to scale up demand.
 
@@
 average service times of primary and secondary vehicles derived from the initial
-simulation model. \textcolor{red}{APPENDIX}~\ref{apx:hyperparameters} shows some initial explorations on
+simulation model. APPENDIX~\ref{apx:hyperparameters} shows some initial explorations on
 the hyper-parameter choices, giving confidence that the chosen parameters, with
@@
 unavailability.
-Looking at the expected survival, that is the value of the MESLMHPHF objective
-function, we see that using an better allocation does increase the survival,
-and that the drop in survival due to demand increases is not as severe when
-using an improved allocation in comparison to the current allocation. We can
-also see that in particular A1 patients are benefiting from the improved
+
+\textcolor{red}{The value of the MESLMHPHF objective function and
+the observed survival, as given by the simulation model, do not match.
+Section~\ref{sec:compare_exp_and_sim} discusses this in more detail.}
+
+\textcolor{red}{As desired, the optimisation heuristic has indeed improved the
+allocation as measured by the MESLMHPHF objective function. In particular A1 patients 
+are benefiting from the improved
 allocations, as these are the patients most sensitive to small fluctuations in
 response time, due to their survival function Equation~\ref{eqn:survival_curve}.
-\textcolor{red}{There is a discrepancy between the expected survival
-measures calculated from the MESLMHPHF objective, and the observed survival from
-the simulation. This is discussed further in the next section.}
+Also, using a better allocation does increase the survival,
+and that the drop in survival due to demand increases is not as severe when
+using an improved allocation in comparison to the current allocation, although
+more numerical evidence would be helpful here. This
+improved robustness 
+not only validates the heuristic but also highlights
+the research contribution of Section~\ref{sec:utilisation} as an approach to
+handle the recursive relationship between an allocation \(x\) and the utilisation
+rates.}
+
 It is worth noting that for the simulation based KPIs (mean response times and
-percent abandoned), there is not too much
+percent abandoned), there is not much
 difference between the currently used allocation and the better allocation,
@@
 
-\textcolor{red}{
 \subsection{Comparing Expected \& Simulated Survival Measures}\label{sec:compare_exp_and_sim}
@@
 times are relatively large in comparison to the mean response times, and so this
-effect would be substantial here.}
-
-\textcolor{red}{
+effect would be substantial here.
+
 The discrepancies are exacerbated by the use of cut-off survival function curves
@@
 model travel times can have a large impact on the survival measures, in this
-case the naive Triangular distribution.}
-
-\textcolor{red}{
+case the naive Triangular distribution.
+
 This leads to a number of avenues of future work to improve this and overcome
@@
 into the objective function and using simulation-optimisation techniques to
-improve runtime efficiency.}
-
-\textcolor{red}{
+improve runtime efficiency.
+
 Due to the naive choice of travel time distributions, including difficulties in
@@
 MESLMHPHF objective from now on.
-}
 
@@
 \caption{MESLMHPHF objective expected survival results.
-\textcolor{red}{(Note that Figures~\ref{fig:results_objective_single} and~\ref{fig:results_objective_multiple} do not have the same y-axis scale in order to keep the data points visible.)}}
+(Note that Figures~\ref{fig:results_objective_single} and~\ref{fig:results_objective_multiple} do not have the same y-axis scale in order to keep the data points visible.)}
 \label{fig:results_objective}
@@
 expected survival, especially those of speciality A1.
-\textcolor{red}{In fact, it can be seen from Figure~\ref{fig:results_a1survival_multiple} that when using multiple vehicle types the survival of A1 patients plateaus as the resource level increases, indicating that the system has reached an upper bound on the survival of A1 patients.
-This is corroborated by considering the survival function of A1 patients given in Equation~\ref{eqn:survival_curve}, where even if all patients were seen immediately a survival of 43.5\% would be expected, and so accounting for even a minimum travel time to the patients an observed plateau value of 32\% could not be improved upon by adding more resource level.}
+In fact, it can be seen from Figure~\ref{fig:results_a1survival_multiple} that when using multiple vehicle types the survival of 
+A1 patients plateaus as the resource level increases, indicating that the system has reached an upper bound on the survival of A1 patients. 
+This is corroborated by considering the survival function of A1 patients given in Equation~\ref{eqn:survival_curve}, where even if all 
+patients were seen immediately a survival of 43.5\% would be expected, and so accounting for even a minimum travel time to the patients an 
+observed plateau value of 32\% could not be improved upon by adding more resource level.
 
@@
 
-\textcolor{red}{In building the models, some assumptions or relaxations were
+In building the models, some assumptions or relaxations were
 made to simplify the model, and future work should investigate the effects of
@@
 Another direction of future work is to find exact methods of solving the
-MESLMHPHF instead of relying on a metaheuristic algorithm.}
+MESLMHPHF instead of relying on a metaheuristic algorithm.
 
@@
 
-\textcolor{red}{
 \section{Full MESLMHPHF Optimisation Model}\label{apx:fullmodel}
@@
 where $\mathcal{L}$ is the resource level.
-}
 

### 2025-01-29 tex/main.tex
include confidence intervals and write up confidence intervals for grids

@@
 in Table~\ref{tbl:current_grid_results}. The simulation was run over half a
-year, with a month warm up time, over eight replications. 
-\textcolor{red}{The variability over the eight replications ...}
+year, with a month warm up time, over five replications. 
+\textcolor{red}{The variability over the replications is very small, and we have included 90\% conficence intervals around the mean response times to demonstrate this.}
 We observe that the
@@
 Ambulance Utilisation & 30.99\% & 36.58\% & 35.79\% \\
-RRV Utilisation & 22.47\% & - & - \\
-Mean Response Time (mins) & 17.69 & 22.29 & 24.05 \\
+RRV Utilisation & 22.47\% & - & - \\[1.5mm]
+\multirow{2}{*}{Mean Response Time (mins)} & 17.69 & 22.29 & 24.05 \\
+ & \textcolor{red}{($\pm$ 0.07)} & \textcolor{red}{($\pm$ 0.06)} & \textcolor{red}{($\pm$ 0.03)} \\[1.5mm]
 StDev Response Time (mins) & 6.98 & 8.02 & 7.94 \\
 MESLMHPHF Objective & 98.34\% & 87.95\% & 85.64\% \\
-Expected A1 Survival & 26.05\% & 13.64\% & 12.23\% \\
+MESLMHPHF A1 Objective & 26.05\% & 13.64\% & 12.23\% \\
 Percent Abandoned & 0.00\% & 0.52\% & 0.93\% \\
@@
 function value and simulation derived KPIs between the current and the optimised
-allocations, for each demand scenario. It can be seen that, as expected, as
+allocations, for each demand scenario\textcolor{red}{, along with 90\% confidence intervals over the trials}. It can be seen that, as expected, as
 demand increases then utilisation of primary and secondary vehicles increase, as
@@
 display the obtained KPIs for each of these allocations. Confidence intervals
-over the eight replications were too small to display on the plots. It can
+over the replications were too small to display on the plots. It can
 immediately be seen that increasing the resource level has a positive effect on

### 2025-01-29 response_2/response.tex

@@
 \begin{itemize}
-
     \item Reviewer 1's comment from the first review:
@@
     \begin{quote}
-``No measures of variance are given for the simulation results in Tables 1
-and 4. Especially since repeated simulation results were conducted, measu-
-res of variance should be included to investigate the statistical significance
-of the results.''
+    ``No measures of variance are given for the simulation results in Tables 1 and 4.
+Especially since repeated simulation results were conducted, measures of variance should be included to investigate the statistical significance of the results.''
     \end{quote}
 
+    \begin{itemize}
         \item Our response to reviewer 1's comment from the first review:
 
-Thank you for the comment, standard deviations of mean response times are now
+        Thank you for the comment, standard deviations of mean response times are no
         given in Table 1 and 4. This is a particularly useful inclusion as the
@@
         \begin{quote}
-Not addressed: The standard deviation of the response time was inclu-
-ded, but no information about the variance of the results over the eight
+        ``Not addressed: The standard deviation of the response time was included, but no information about the variance of the results over the eight
         replications was given. To obtain knowledge about the reliability of the
@@
         simulation results differed between the eight replications mentioned on
-page 17.
+        page 17.''
         \end{quote}
 
-A short discussion about the variability over the repetitions (including
-variance) has now been included. 
-% TODO VK: We need to do this, I believe it is only Geraint who has the data.
+        Thank you for the comment.
+Firstly, this highlighted a small error in the paper, five repetitions were conducted rather than eight; this is found in the source code given in \url{https://github.com/MarkTuson/ambulance_simulation}.
+In the updated manuscript 90\% confidence intervals for some KPIs have now been included in Tables 1 and 4, namely the simulated survival and mean response times.
+A short comment about the variability over the repetitions has now been included.
+The variance over the repetitions is very small, so small that as mentioned in the original draft of the manuscript, confidence intervals were not able to be visualised on the plot in Figures 10 onwards.
+    \end{itemize}
 
@@
     \begin{quote}
-``Unclear, why the subchapter 4.1 on survival functions is necessary, sin-
-ce traditional coverage targets are used in the case study and the model
-formulation does not depend on characteristics of the survival functions.
-While survival functions are an important performance metric in the re-
-viewers’ opinion, it is not clear why the reader needs this information to
-follow the remainder of the pape.''
+    ``Unclear, why the subchapter 4.1 on survival functions is necessary, since traditional coverage targets are used in the case study and the model formulation does not depend on characteristics of the survival functions.
+While survival functions are an important performance metric in the reviewers' opinion, it is not clear why the reader needs this information to follow the remainder of the pape.''
     \end{quote}
 
+    \begin{itemize}
         \item Our response to reviewer 1's comment from the first review:
 
-We disagree with the reviewer here. In the case study, traditional targets such
-a mean response times are reported alongside survival measures, based on
-the survival functions. However, a key part of the paper is the
-optimisation process, and all scenarios compared are `optimised'
-scenarios that result from this process. The objective function of the
-optimisation process maximises survival, based on survival functions.
-This is evnidenced in Equations 6 and 7, which are component parts of
-the objective function given in Equation 5. Therefore the survival
-functions are central to the work presented.
+        We disagree with the reviewer here.
+In the case study, traditional targets such a mean response times are reported alongside survival measures, based on the survival functions.
+However, a key part of the paper is the optimisation process, and all scenarios compared are `optimised' scenarios that result from this process.
+The objective function of the optimisation process maximises survival, based on survival functions.
+This is evnidenced in Equations 6 and 7, which are component parts of the objective function given in Equation 5.
+Therefore the survival functions are central to the work presented.
 
@@
         \begin{quote}
-''Not addressed: It was expanded upon why survival functions can be sui-
-table performance metrics. In the case study however, coverage targets
-were used (8 min target for A1 patients, 15 min target for A2 patients
-and 60 min target for B patients). It is therefore still unclear, why the
-reader needs the information in chapter 4.1. Additionally, there is no dis-
-cussion as to why coverage targets were chosen in the case study, after
-it was argued, that survival functions can solve some drawbacks of using
-coverage targets.``
+        ``Not addressed: It was expanded upon why survival functions can be suitable performance metrics.
+In the case study however, coverage targets were used (8 min target for A1 patients, 15 min target for A2 patients and 60 min target for B patients).
+It is therefore still unclear, why the reader needs the information in chapter 4.1.
+Additionally, there is no discussion as to why coverage targets were chosen in the case study, after it was argued, that survival functions can solve some drawbacks of using coverage targets.``
         \end{quote}
@@
         \begin{quote}
-    ``
-There is contradicting information about which survival functions were
-used in the case study. On page 15 and 16, response time targets for the
-different categories are listed. This implies to the reader, that step-wise
-survival functions, in other words coverage targets, were used in the case
-study. On page 19, line 44 it is stated, that Equation 3 was used as the
-survival function for A1 patients. It needs to be clarified, which survival
-functions were used in the MESLMHPHF in the case study both for the
-optimization metaheuristic and the evaluation of the results.
-    ''
+        ``There is contradicting information about which survival functions were used in the case study.
+On page 15 and 16, response time targets for the different categories are listed.
+This implies to the reader, that step-wise survival functions, in other words coverage targets, were used in the case study.
+On page 19, line 44 it is stated, that Equation 3 was used as the survival function for A1 patients.
+It needs to be clarified, which survival functions were used in the MESLMHPHF in the case study both for the optimization metaheuristic and the evaluation of the results.''
         \end{quote}
@@
 
-We have added a sentence to the text to clarify that despite the policy target
-of 8 minutes for the A1 patients, we have in fact used the survival function of
-equation (3) for A1 patients. This was not explicit and we appreciate the
-reviewer insisting this point.
+        We have added a sentence to the text to clarify that despite the policy target of 8 minutes for the A1 patients, we have in fact used the survival function of equation (3) for A1 patients.
+This was not explicit and we appreciate the reviewer insisting this point.
+
+    \end{itemize}
 
@@
     \begin{quote}
-``The simulated survival probabilities in Table 4 show, that the survival for
-A1 patients was not improved in any of the four scenarios in the simu-
-lation. As the simulation model is a more realistic model of the actual
-1
-EMS system in comparison to the MESLMHPHF, the simulated survi-
-val probabilities carry more weight in interpreting the results compared
-to the expected survival probabilities calculated with the MESLMHPHF.
-On page 19, line 42 - 44 however it is stated, that ‘in particular A1 pati-
-ents are benefitting from the improved allocations’. It is unclear, how the
-authors draw this conclusion, when the simulation results show worse or
-equal survival probabilities for A1 patients in all scenarios.''
+    ``The simulated survival probabilities in Table 4 show, that the survival for A1 patients was not improved in any of the four scenarios in the simulation.
+As the simulation model is a more realistic model of the actual EMS system in comparison to the MESLMHPHF, the simulated survival probabilities carry more weight in interpreting the results compared to the expected survival probabilities calculated with the MESLMHPHF.
+On page 19, line 42 - 44 however it is stated, that ‘in particular A1 patients are benefitting from the improved allocations’.
+It is unclear, how the authors draw this conclusion, when the simulation results show worse or equal survival probabilities for A1 patients in all scenarios.''
     \end{quote}
 
-The reviewer raises a good point here. We have modified the text to hopefully
-clarify that:
+    The reviewer raises a good point here. We have modified the text to hopefully clarify that:
 
     \begin{itemize}
-    \item The sentence regard A1 patients meant to reflect more on the
-        efficiency and accuracy of the optimisation heuristic which we find is
-        one of the main contributions of this paper (specifically the novel
-        approach used to deal with the circular nature of utilisation and
-        allocation).
-    \item The optimised allocation give more robust allocations under increasing
-        demand although we admit that there is not enough numerical evidence to
-        be sure of this effect.
+        \item The sentence regard A1 patients meant to reflect more on the efficiency and accuracy of the optimisation heuristic which we find is one of the main contributions of this paper (specifically the novel approach used to deal with the circular nature of utilisation and allocation).
+        \item The optimised allocation give more robust allocations under increasing demand although we admit that there is not enough numerical evidence to be sure of this effect.
     \end{itemize}
 
-Furthermore we have more explicitly referred to our explanation about the cause
-of this discrepancy.
+    Furthermore we have more explicitly referred to our explanation about the cause of this discrepancy.
 
-Fundamentally, again, the reviewer raises a valid and interesting avenue for investigation and
-one that we feel
-could not be precisely considered without the work in this paper. We hope to
-tackle building a more accurate objective function, that better captures the
-stochastic effects, that could be then optimised with our novel algorithm. This,
-does not however fit in this paper. If the reviewer would like us to make that
-more explicit and clear in the text we would be happy to follow any guidance
-they feel is necessary.
+    Fundamentally, again, the reviewer raises a valid and interesting avenue for investigation and one that we feel could not be precisely considered without the work in this paper.
+We hope to tackle building a more accurate objective function, that better captures the stochastic effects, that could be then optimised with our novel algorithm.
+This, does not however fit in this paper.
+If the reviewer would like us to make that more explicit and clear in the text we would be happy to follow any guidance they feel is necessary.
 
   \item Reviewer 1's comment:
+    \begin{quote}
+    ``Equations 8 and 9: It is stated, that \(\beta_{pa_1a_2}\) indicates, that a vehicle from \(a_1\) can reach \(p\) quicker, than a vehicle from \(a_2\).
+The equations 8 and 9 however use \(\leq\) and not \(<\) , which is not consistent with the description.
+Has been fixed for \(\beta_{pa_1a_2}\) in Equation 8, but not for \(R_{pa_1a_2}\) in Equation 9.''
+    \end{quote}
 
-    \begin{quote}
-        ``      Equations 8 and 9: It is stated, that \(\beta_{pa_1a_2}\) indicates, that a vehicle from
-        \(a_1\) can reach \(p\) quicker, than a vehicle from \(a_2\). The equations 8 and 9
-        however use \(\leq\) and not \(<\) , which is not consistent with the description.
-Has been fixed for \(\beta_{pa_1a_2}\) in Equation 8, but not for \(R_{pa_1a_2}\) in Equation
-9. ''
-    \end{quote}
-    This has been fixed.
-    % TODO Check with Geraint that this is correct.
+    We have reverted both to be $\leq$, to correspond with the source code.
+We have added and explanatory note to the text explaining the edge cases.
+In Equation 8 there is no difference between using $<$ and $\leq$, as we assume no equidistant ambulance stations; while in Equation 9 using $\leq$ is necessary as primary vehicles take priority over secondary vehicles in the event of a tie.
 

### 2025-01-29 tex/main.tex

@@
 it is assumed that no two distinct ambulance locations are exactly equidistant
-from a pickup location. Note that here these parameters are defined using
+from a pickup location.
+\textcolor{red}{In both Equation~\ref{eqn:beta} and~\ref{eqn:R}, $\leq$ are used rather than $<$; in Equation~\ref{eqn:beta} will be equivalent, as we assume no two ambulance locations are equidistant; while in Equation~\ref{eqn:R}, in the case of a primary and secondary vehicle having equal travel times, then the primary vehicle will have priority.}
+Note that here these parameters are defined using
 travel times, but can be generalised to account for any type of preference,
@@
     0 & \text{if } a_1 = a_2\\
-    1 & \text{if } b_{p a_1} < b_{p a_2}\\
+    1 & \text{if } b_{p a_1} \leq b_{p a_2}\\
     0 & \text{otherwise.}
